### 2018-12-04 README.Rmd
Generate Readme + add notes

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# fishcom

The goal of fishcom is to get all the informations concerning the ECOSTAB postdoc
project in one place including reproducible code, notes, eventually data and
manuscript.

## How to

My notes are available in the introductory vignette
[intro.html](https://github.com/alaindanet/fishcom/blob/master/vignettes/intro.html)