### 2019-11-13 report/06-pressure.Rmd

## Pressure

```{r}
myload(temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
myload(temporal_network_metrics, dir = dest_dir)
myload(community_metrics, temporal_community_metrics, synchrony,
temporal_press_polluants, press_metrics, op_analysis,
  dir = data_common)
```

```{r}
var_values <-  temporal_press_polluants %>%
  select(id, category, press) %>%
  spread(category, press) %>% 
  ungroup()

m <- cor(na.omit(select(var_values, -id)))

as.data.frame(round(m, 2)) %>%
  mutate_all(list(~cell_spec(., "html", color = ifelse(abs(.) > .65 & . != 1, "red",
	  "black")))) %>%
  mutate(variable = colnames(.)) %>%
  select(variable, everything()) %>%
  kable(format = "html", escape = F)
```

```{r press-pca, fig.cap = "PCA of the different press category."}
res.pca <- dudi.pca(na.omit(select(ungroup(var_values), -id)),
  scannf = FALSE,   # Hide scree plot
  nf = 5            # Number of components kept in the results
)

eig_plot <- fviz_eig(res.pca)

pca_plot <- lapply(list(c(1,2), c(1, 3), c(3, 2)), function (axes) {
  fviz_pca_var(res.pca,
    axes = axes,
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE     # Avoid text overlapping
  )
})

plot_grid(eig_plot, pca_plot[[1]], pca_plot[[2]], pca_plot[[3]], labels = "AUTO")
```

The PCA performed poorly (Fig. \@ref(fig:press-pca)A), the two first axis explaining
less than 40% of the variance. However, the first axis showed that the DBO,
phosphorus, nitrates, temperature and herbicides co-vary strongly (Fig.
\@ref(fig:press-pca)B), DBO and phosphorus being the more strongly related.
Interestingly, insecticides concentration was orthogonal with DBO (meaning no
correlation?) (Fig. \@ref(fig:press-pca)C and D). Fungicides and herbicides seems to
covary positively whereas insecticides seemed to covary negatively with the
formers (Fig. \@ref(fig:press-pca)B and C). We also found that the percentage of
dissolved was negatively linked with temperature and DBO (Fig.
\@ref(fig:press-pca)C and D), which makes sense.

### Effects on stability

```{r}
cat_to_rm <- c("mes", "disolved_oxygen",
  "flow", "industry")
temporal_press_polluants %<>%
  filter(!category %in% c(
      "matieres azotees", "other", "matieres phosphorees", "matieres
      organiques"))
```

```{r}
biomass_com <- temporal_community_metrics %>%
  select(station, biomass_stab, richness_med, betadiv)
biomass_com %<>%
  left_join(rename(temporal_press_polluants, station = id), by = "station")
biomass_com %<>% na.omit()

pca_press_data <- biomass_com %>%
  select(station, richness_med, category, press) %>%
  filter(!category %in% cat_to_rm) %>%
  spread(category, press)
```

```{r, fig.dim = c(4, 4)}
res.pca <- dudi.pca(na.omit(select(pca_press_data, -station)),
  scannf = FALSE,   # Hide scree plot
  nf = 5            # Number of components kept in the results
)

eig_plot <- fviz_eig(res.pca)
```

```{r pca-press2, fig.dim = c(10, 7), fig.cap = "PCA with the candidate pressure"}
pca_plot <- lapply(list(c(1,2), c(1, 3), c(1, 4), c(3, 2), c(4, 2), c(4, 3)), function (axes) {
  fviz_pca_var(res.pca,
    axes = axes,
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE     # Avoid text overlapping
  )
})
plot_grid(eig_plot, pca_plot[[1]], pca_plot[[2]], pca_plot[[3]], pca_plot[[4]],
  pca_plot[[5]], labels = "AUTO")
```

## SEM

```{r}
mod <- lm(betadiv ~ log(richness_med), data = temporal_community_metrics)
temporal_community_metrics$betadiv_corrected <- resid(mod)
```

```{r}
biomass_com <- temporal_community_metrics %>%
  select(station, biomass_stab, richness_med, betadiv, simpson_med, pielou_med,
  betadiv_corrected)
biomass_com %<>%
  left_join(rename(temporal_press_polluants, station = id), by = "station")
biomass_com %<>% na.omit()
biomass_com %<>% 
  select(-cv_press, -press) %>%
  spread(category, press_med) %>%
  left_join(synchrony)
net_com <- temporal_network_metrics %>%
  select(station, connectance_corrected_med, modularity_corrected_med,
    mean_troph_level_corrected_med, w_trph_lvl_avg_med)

dsem <- biomass_com  %>%
  mutate(log_rich = log10(richness_med)) %>%
  left_join(net_com, by = "station") %>%
  na.omit()
```

### All

```{r}
library(lavaan)
options(width = 100)
lav_data <- dsem %>%
  mutate(log_cv_sp = log10(cv_sp),
    log_biomass_stab = log10(biomass_stab)
  )
lav_formula <- " 
  synchrony ~ log_rich + pielou_med + temperature + connectance_corrected_med + w_trph_lvl_avg_med + betadiv_corrected
  log_cv_sp ~ log_rich + pielou_med + temperature + connectance_corrected_med + w_trph_lvl_avg_med + betadiv_corrected
  log_biomass_stab ~ log_rich + pielou_med + connectance_corrected_med + w_trph_lvl_avg_med + betadiv_corrected
  log_rich ~ temperature + DBO + fungicides + herbicides + insecticides
  betadiv_corrected ~ temperature + DBO + fungicides + herbicides + insecticides
  connectance_corrected_med ~ temperature + DBO + fungicides + herbicides + insecticides
  pielou_med ~ temperature + DBO + fungicides + herbicides + insecticides
  w_trph_lvl_avg_med ~ temperature + DBO + fungicides + herbicides + insecticides
  DBO ~ nitrates + phosphore + fungicides + herbicides + insecticides
  "
lav_mod <- sem(lav_formula, lav_data)
summary(lav_mod, standardized = TRUE)
```

```{r, fig.dim = c(10, 7)}
ggsem <- function(fit, layout = "sugiyama") {

  # Extract standardized parameters
  params <- lavaan::standardizedSolution(fit)

  # Edge properties
  param_edges <- params %>% 
    filter(op %in% c("=~", "~", "~~"), lhs != rhs, pvalue < .05) %>%
    transmute(to = lhs,
              from = rhs,
              val = est.std,
              type = dplyr::case_when(
                op == "=~" ~ "loading",
                op == "~"  ~ "regression",
                op == "~~" ~ "correlation",
                TRUE ~ NA_character_))

  # Identify latent variables for nodes
  latent_nodes <- param_edges %>% 
    filter(type == "loading") %>% 
    distinct(to) %>% 
    transmute(metric = to, latent = TRUE)

  # Node properties
  param_nodes <- params %>% 
    filter(lhs == rhs) %>% 
    transmute(metric = lhs, e = est.std) %>% 
    left_join(latent_nodes) %>% 
    mutate(latent = if_else(is.na(latent), FALSE, latent))

  # Complete Graph Object
  param_graph <- tidygraph::tbl_graph(param_nodes, param_edges)

  # Plot
  ggraph(param_graph, layout = layout) +
    # Latent factor Nodes
    geom_node_point(aes(alpha = as.numeric(latent)),
                                     shape = 16, size = 5) +
    geom_node_point(aes(alpha = as.numeric(latent)),
                                     shape = 16, size = 4, color = "white") +
    # Observed Nodes
    geom_node_point(aes(alpha = as.numeric(!latent)),
                                     shape = 15, size = 5) +
    geom_node_point(aes(alpha = as.numeric(!latent)),
                                     shape = 15, size = 4, color = "white") +
    # Regression Paths (and text)
    geom_edge_link(aes(color = val, label = round(val, 2),
                       alpha = as.numeric(type == "regression")),
                   linetype = 1, angle_calc = "along", vjust = -.5,
                   arrow = arrow(20, unit(.3, "cm"), type = "closed")) +
    # Factor Loadings (no text)
    geom_edge_link(aes(color = val, alpha = as.numeric(type == "loading")),
                   linetype = 3, angle_calc = "along",
                   arrow = arrow(20, unit(.3, "cm"), ends = "first", type = "closed")) +
    # Correlation Paths (no text)
    geom_edge_link(aes(color = val, alpha = as.numeric(type == "correlation")),
                   linetype = 2, angle_calc = "along",
                   arrow = arrow(20, unit(.3, "cm"), type = "closed", ends = "both")) +
    # Node names
    geom_node_text(aes(label = metric),
                   nudge_y = .25, hjust = "inward") +
    # Node residual error
    geom_node_text(aes(label = sprintf("%.2f", e)),
                   nudge_y = -.1, size = 3) +
    # Scales and themes
    scale_alpha(guide = FALSE, range = c(0, 1)) +
    scale_edge_alpha(guide = FALSE, range = c(0, 1)) +
    scale_edge_colour_gradient2(guide = FALSE, low = "red", mid = "darkgray", high = "green") +
    scale_edge_linetype(guide = FALSE) +
    scale_size(guide = FALSE) +
    theme_graph()
}
ggsem(lav_mod)

```

### By basin

```{r}

myload(the_8_hydrologic_basin, station_analysis, dir = data_common)

station_analysis %<>%
  st_transform(crs = 2154)
the_8_hydrologic_basin %<>%
  mutate(NomDistric = ifelse(CdBassinDC == "B2", "SAMBRE", NomDistric))

station_basin <- st_intersects(station_analysis, the_8_hydrologic_basin)
station_analysis$basin <- purrr::map_chr(station_basin, function(x){
  if (length(x) == 0) {
   return(NA) 
  }
  the_8_hydrologic_basin[["NomDistric"]][x]
})
# Filter NA
station_analysis %<>%
  filter(!is.na(basin)) %>%
  mutate(station = id) %>%
  select(station, basin)
st_geometry(station_analysis) <- NULL

# How many station by basin:
nb_st_basin <- station_analysis %>%
  group_by(basin) %>%
  summarise(nobs = n())
```

```{r}
lav_data %<>%
  left_join(station_analysis, by = "station")

lav_data_test <- filter(lav_data, basin %in% nb_st_basin[nb_st_basin$nobs > 61, ]$basin)

# Since the group 1 and 3 are two small, I cut two basin 
unique(lav_data$basin)
lav_mod_basin <- sem(lav_formula, lav_data_test, group = "basin")
summary(lav_mod_basin, standardized = TRUE)
```

```{r, fig.dim = c(14, 10)}
lav_basin_list <- lav_data_test %>%
  group_by(basin) %>%
  nest() %>%
  mutate(lav_mod = map(data, function(.data){ sem(lav_formula, .data)}),
      plot_lav = map(lav_mod, function(lav){ ggsem(lav)})
      )
plot_grid(plotlist = lav_basin_list$plot_lav)
```

### Model with only community structure and stability

```{r, fig.dim = c(14, 10)}
lav_formula_str_stab <- " 
  synchrony ~ log_rich + pielou_med + connectance_corrected_med + w_trph_lvl_avg_med + betadiv_corrected
  log_cv_sp ~ log_rich + pielou_med + connectance_corrected_med + w_trph_lvl_avg_med + betadiv_corrected
  log_biomass_stab ~ log_rich + pielou_med + connectance_corrected_med + w_trph_lvl_avg_med + betadiv_corrected
  "

lav_basin_str_stab_list <- lav_data %>%
  group_by(basin) %>%
  nest() %>%
  mutate(lav_mod = map(data, function(.data){
      out <- tryCatch(
	{sem(lav_formula_str_stab, .data)},
	error = function(cond) {return(NA)},
	warning = function(cond) {return(NA)}
      )
      return(out)
		       }),
      plot_lav = map(lav_mod, function(lav){
      out <- tryCatch(
	{ggsem(lav)},
	error = function(cond) {return(NA)},
	warning = function(cond) {return(NA)}
      )
      return(out)
})
      )

lav_basin_str_stab_list %<>%
  mutate(class_plot = map_lgl(plot_lav, is.logical))%>%
  filter(!class_plot)
plot_grid(plotlist = lav_basin_str_stab_list$plot_lav)
```