### 2020-03-04 report/08-infer-interactions.Rmd
Inverse PCA axis

## Infer species association

The goal is to infer species associations thanks to `PLNmodels` R package.

```{r}
myload(temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
myload(temporal_network_metrics, dir = dest_dir)
myload(network_analysis, dir = mypath("data", "classes"))
myload(community_analysis, dir = data_common)
```

```{r get-fish-network-abun}
fish_abun <- network_analysis %>%
  unnest(composition) %>%
  select(opcod, sp_class, nind) %>%
  filter(!is.na(opcod))
```

```{r get-opcod-covariate}
myload(habitat_pressure, dir = mypath("report"))
myload(press_metrics, flow_temp_metrics, dir = data_common)
myload(env_analysis, geo_station, dir = data_common)
myload(op_analysis, dir = data_common)

# DBO:
press <- press_metrics %>%
  filter(category %in% c("DBO")) %>%
  rename(station = id) %>%
  spread(category, press)
# temperature and flow 
flow_temp <- flow_temp_metrics %>%
  ungroup() %>%
  rename(station = id) %>%
  spread(category, press)
# River desc:
env <- env_analysis %>%
  select(opcod, width_river, avg_depth_station)
# Alt, d_source
position <- geo_station %>%
  select(station, alt, d_source)

# Collect data:
## Fish operation:
op <- op_analysis %>%
  select(opcod, station, year, surface, length_sourced) %>%
  left_join(env, by = c("opcod"))

## covariate table:
covariates <- press %>% 
  left_join(flow_temp, by = c("station", "year")) %>%
  left_join(position, by = c("station")) %>%
  left_join(
    select(op, opcod, station, year, width_river, avg_depth_station), by = c("station", "year")) %>%
  select(-station, -year) %>%
  filter(!is.na(opcod))

```

```{r}

# Sanitize opcod:
map_dbl(list(fish_abun, covariates), ~length(unique(.x$opcod))) 
fish_ok <- fish_abun %>%
  filter(opcod %in% unique(covariates$opcod)) %>%
  arrange(opcod)
cov_mat <- covariates %>%
  filter(opcod %in% unique(fish_abun$opcod)) %>%
  arrange(opcod) %>%
  as.data.frame

rownames(cov_mat) <- cov_mat[, "opcod"] 
cov_mat  <- cov_mat[, !colnames(cov_mat) %in% "opcod"]

# remove opcod and build abundance matrix
mat_abun <- fish_abun %>%
  spread(sp_class, nind) %>%
  as.data.frame

rownames(mat_abun) <- mat_abun[, "opcod"] 
mat_abun <- mat_abun[, !colnames(mat_abun) %in% "opcod"]

save_dir <- mypath("report", "stephane")
if (!dir.exists(save_dir)) {
  dir.create(save_dir)
}
mysave(mat_abun, cov_mat, dir = save_dir)
```

```{r}
library(PLNmodels)
fish <- prepare_data(
  counts = mat_abun, 
  covariates = cov_mat)
network_models <- PLNnetwork(Abundance ~ 1 + offset(log(Offset)), data = fish)
plot(network_models, "diagnostic")
plot(network_models)
network_models
model_StARS <- getBestModel(network_models, "StARS")

my_graph <- plot(model_StARS, plot = FALSE)
plot(model_StARS)
plot(model_StARS, type = "support", output = "corrplot")
```

```{r}
data(trichoptera)
trichoptera <- prepare_data(trichoptera$Abundance, trichoptera$Covariate)
network_models <- PLNnetwork(Abundance ~ 1 + offset(log(Offset)), data = trichoptera)
plot(network_models)
model_StARS <- getBestModel(network_models, "StARS")
my_graph <- plot(model_StARS, plot = FALSE)
plot(model_StARS)
```