### 2020-10-29 manuscript/bef_stability/appendix.Rmd
Improve appendix clarity

---
title: Appendix
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

```{r, echo = FALSE, message = FALSE, collapse = FALSE}

mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

load(file = mypath("manuscript", "bef_stability", "result", "workspace.rda"))

knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))
source(mypath("R", "cleaning_methods.R"))

theme_set(theme_alain())
```

```{r, echo = FALSE}
package <- c("openSTARS", "SSN", "nlme", "piecewiseSEM")
#purrr::map(package, ~ toBibtex(citation(.x)))

#toBibtex(citation())

#print(cm, bibtex = TRUE)
# Load manuscript results
```

# Data collection and preparation

## Fishing

### Protocol

The fish community sampling took place in a variety of stream size, shape and
depth. This diversity constraints the type of protocol that can take place.  For
the small streams (depth < 0.7m across the entire width), the complete protocol
is used. It consists in placing anodes every X meters across the width of the
river. The fishermen made from one to three row
of fishing. We kept only the first row for analysis to homogenize the sampling
effort. For deeper and wider streams, the sampling was partial over the bank and
the channel and performed by boat and by foot (along the bank). We kept only
sampling done over the bank because channel sampling was not always done. Two
types of partial protocols takes place: partial over bank and partial by points.
The former is continuous while the latter is discrete. Furthermore, in partial
by point protocol the channel can be sampled as well. We kept only the partial
by point fish sampling which all the sampling point were done over the bank to
homogenize the sampling. Then, we merged partial over bank and partial over
point protocols. Some sampling stations had mix characteristics in which mix of
partial and complete protocols were applied depending of the depth of the river
at the time of sampling.

Heterogeneity in the sampling period may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in sampling time, only one sampling per year was retained if there
was two in a given year, the one with with the most species and then individual
collected. Finally, most of the fish sampling occurred in September (Median:
mid-August) and the median standard deviation of the sampling month was inferior
to one (0.7) within station. Additionally, we removed sampling events, in a
given station, which the sampled length along the stream was outside the median
length sampled more or less 30%. Finally, we standardized the biomass record
and species richness by the sampled surface to get rid of sampling effort
effects.

### Batches definition and body size measurement

During the electric fishing, and before that fish body size are measured, they
are regrouped by batches. Each batches contains one species but there can be
several batches by species. Body size assessment are different according to the
batches. There are four types of batches: N, I, S/L and G. The batches N and
I contains the biggest individuals, there are normally less numerous and then
the body size of the individuals of those batches is recorded. The batches S/L
contains fishes of intermediate sizes. Inside each S/L batches, the operator
tries to make batches as homogeneous as possible in body size. Then, the body
size of a subsample of X individuals are recorded. The G batches contain
individuals of low body size, with care that body size is as homogeneous as
possible in the batches. Then body size of the shortest and the longest
individuals are recorded.

## Body size inference

```{r}
myload(fish_length, lot_id_opcod, dir = mypath("data-raw", "fishing_op_build"))

fish_length %<>%
  left_join(lot_id_opcod) %>%
  rename(length = fish) %>%
  select(opcod, species, length)

length_sem_analysis <- fish_length %>% 
  filter(opcod %in% op_analysis_bbb$opcod)
rm(fish_length, lot_id_opcod)
nrow_fish_length <- nrow(length_sem_analysis)
na_length <- sum(is.na(length_sem_analysis$length))
```

The instructions for body size inference were guided by the recommendations of
the OFB agency, which has defined the sampling protocols, supervised the data
collection and performed analysis for multiple purpose including stream
management.

```{r}
body_size_caption <- paste0(
  "Body size inference of the fishes. ",
  "(1) Sampled fishes were attributed to batch type according to their size and
  the number of individuals. A batch contains one species only. (2) In the batch
  type N and I, the body size of all the individuals was recorded. In the
  batches, S/L, the body size of a subsample of the individuals was recorded.
  Finally, only the body size of the littlest and of the tallest individual was
  recorded in the G batches. ",
  "Then, (3) we inferred the body size distribution in the S/L and G batches by
  assuming that it followed a normal distribution. The mean and the standard
  deviation were estimated with the measured subsample."
) 
```

```{r p-body-size, fig.cap=body_size_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "length_inference.pdf"))
```

With the information contained in the batch record, we inferred the body size of
the fishes in the community. No inference were needed for the N and I batches
because the body size of the individuals were recorded. The body size ($l$) of the
fishes in the S/L and G were inferred under the assumption that its distribution
followed a normal law ($l ~ N(\mu , \sigma)$).
For a S/L batch, the mean $\mu$ and the standard deviation $\sigma$ were simply
the average and the standard deviation of the body size of the subsample
recorded. For a G batch, the average was approximated by 
$\hat{\mu} = (l_{min} + l_{max}) 2^{-1}$ and the standard deviation by
$\hat{\sigma} = (l_{min} + l_{max}) 4^{-1}$. Together with the estimations of
the mean and the standard deviation, the number of individuals in a batch and
assumption of a normal distribution, we simulated the body size distribution in
S/L and G batches. To avoid simulation of unrealistic body size, we used a
truncated normal distribution bounded at [0.05; .95] for the density of
probability of the S/L batches, and bounded at [$l_{min}$; $l_{max}$] for the G
batches. Body size distribution were generated using the R function
`truncdist::rtrunc`.

We removed batches that did not satisfied the following conditions. First, we
removed batches which the total number of individuals was not recorded. In lot
"G", we removed batches were  $l_{min} \geq l_{max}$ or if at least one of the
measurement was missing. Additionally, G batches were removed when number of
individuals fell below five, as G batches normally contained more individuals.
In the S/L lot, we removed batches were the number of individuals in the
recorded individuals fell below 10. In sum, those sanity check allowed to
exclude potentially problematic batches. Finally, in the dataset used in the
analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

## Community data

```{r}
source(mypath("R", "cleaning_methods.R"))
removed_ind <- filter(length_sem_analysis, species %in% sp_to_remove())
removed_sp <- unique(removed_ind$species)
```

We removed rare fish species, species from lake, crayfish species, and migratory
species. In total, we removed `r nrow(removed_ind)` over 
`r nrow(length_sem_analysis)` individuals and `r length(removed_sp)`.

During fish sampling, some species are identified based on morphotype but they
belong to the same species based on fecundation criteria.
We merged those morphotype species into one species. There was
`r length(unique(names(sp_to_replace())))` to merge which belonged to 
`r length(unique(sp_to_replace()))` species.

### Metaweb

#### Ontogenic diet

Diet information were obtained from fishbase.org and literature [REF willem?]. The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had as strict piscivorous
life stage. Fish species were largely omnivorous.

#### Define trophic interactions

```{r}
trophic_building_caption <- paste0(
  "Trophic species"
) 
```

```{r p-net-inf, fig.cap=trophic_building_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "network_inference.pdf"))
```

```{r}
myload(metaweb_analysis, dir = mypath("data"))
species_in_metaweb <- metaweb_analysis$species

length_sem_analysis %<>%
  filter(!species %in% sp_to_remove()) %>%
  mutate(
    species = str_replace_all(species, sp_to_replace())
  )

removed_ind_metaweb <- length_sem_analysis %>%
  filter(!species %in% species_in_metaweb)
removed_sp_metaweb <- unique(removed_ind_metaweb$species)
```

The metaweb described the potential trophic interactions between the
($`r ncol(metaweb_analysis$metaweb) - length(metaweb_analysis$resource)`$) trophic species.
We had seven resource nodes which were present in the diet of the trophic
species namely detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos.

The metaweb building was based on trait matching, so it necessitates to have
information about their diet. `r length(removed_sp_metaweb)` species were
removed because information about their diet was missing. It represented  
`r nrow(removed_ind_metaweb)` individuals over `r nrow(length_sem_analysis)` 
individuals.

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
resource nodes were determined by literature. For example, phytobenthos fed on
detritus, while zooplankton fed on zooplankton and phytoplankton. Fish-fish
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies [@brose_predator_2019; @poisot_structure_2016;
@schneider_body_2012].  For a given trophic species, we first determined if its
size class end up in a life stage containing piscivorous diet, i.e. if there was
an overlap between them. We were more liberal for piscivory determination to not
miss potential fish-fish trophic interactions which were the main interactions
in our community. If the trophic species was piscivorous, we determined its prey
size range from 3% to 45% of the median of its size range. A trophic interaction
was set for each trophic species whose median of its size class fell into the
prey range of the predator trophic species. To sum up, the metaweb was obtained
with allometric feeding interactions (1) between trophic species, (2) between
trophic species and resource nodes and (3) between resource nodes.

#### Local trophic network

The local trophic network (i.e. one network for each fish community
sampling) was obtained by determining the trophic species of each fish
individual in the community and interactions were set according to the metaweb.
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species. The inferred interactions represent potential interactions
(i.e. not measured interactions) and are binary (i.e. presence or absence of
feeding interactions).

## Environmental variables

The procedure to obtain altitude, slope, width depth of the streams are fully
explained in main text. Because those measurements were collected during the
fishing protocol, they were easier to manage in the analysis. Then, we detailed
in the following sections how water temperature, flow and BOD were obtained.

### Preparation

Water temperature, flow and BOD (Biological Oxygen Demand) were obtained from the
[naiades database](http://www.naiades.eaufrance.fr/).
Water temperature, flow and BOD were respectively collected at a hourly, daily
and monthly frequency. Water temperature and flow are recorded by automated
sensors while BOD is collected manually because it necessitates analysis in a
laboratory. BOD data collection is coordinated by the Agence de l'eau which
realized the analysis itself or order it to a company.
It is crucial to make clear that those variables are collected at different
places than the fish samplings.

```{r}
pressure_cap <- paste0(
  "Map of environmental variables obtained for the 99 sites used in the analysis."
)
```

```{r, fig.height = 20/2.54}
pca_pressure <- habitat_press
myload(fourier, dir = mypath("data-raw", "polluants"))
myload(flow_temp_metrics, station_analysis, the_8_hydrologic_basin, dir = mypath("data"))

sf_pressure <- station_analysis %>% 
  filter(id %in% op_analysis_bbb$station) %>%
  select(id) %>%
  left_join(rename(pca_pressure, id = station)) %>%
  na.omit()

p <- ggplot() +
  geom_sf(data = the_8_hydrologic_basin)

var_to_plot <- colnames(pca_rotated$normal$tab)

var_replacement <- c(
  "slope" = "Slope",
  "alt" = "Altitude",
  "d_source" = "Distance from source",
  "strahler" = "Strahler order",
  "width_river_mean" = "Average stream width",
  "avg_depth_station_mean" = "Average stream depth",
  "width_river_cv" = "CV of stream width",
  "avg_depth_station_cv" = "CV of stream depth",
  "DBO_med" = "Average BOD",
  "flow_med" = "Average flow",
  "temperature_med" = "Average temperature",
  "DBO_cv" = "CV of BOD",
  "flow_cv" = "CV of flow",
  "temperature_cv" = "CV of temperature"
)

map_pressure <- map(var_to_plot,
~p +
  geom_sf(data = sf_pressure, aes_string(color = .x), size = .5) +
  viridis::scale_color_viridis() +
  labs(title = str_replace_all(.x, var_replacement)) +
  theme(legend.title = element_blank(),
    text = element_text(size = 8),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
 guides(fill = guide_colourbar(barwidth = 0.5, barheight = 10))
)

plot_grid(plotlist = map_pressure, ncol = 3)
```

The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each recording
station, we applied a moving window averaging (more details in supplementary
material) to get rid of seasonal effects. For temperature, the moving was performed on daily
average temperature. Then we took the annual average of each year and each
variable (i.e. one value by year and by station).

Several characteristics of the streams make the interpolation more complicated
than along terrestrial surface. Interpolations in terrestrial areas are roughly
based on euclidean distance. Euclidean distance is more complicated to consider
in streams because it had the follow the path of the streams. Furthermore,
two streams can be close by but not connected and the value of their variables
could be not related or a only a few. For those reason, we used stream spatial
network methods to interpolate temperature, flow and BOD.

We reconstructed the stream network and the connection between sites with the
`openSTARS` R package. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(IGN)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network [@pella_reseau_2012].

The procedure was done by hydrographic basin because of convergence problem
when the interpolation was done at the France scale. We had to make a trade-off
between the number of data (to keep enough high to perform the interpolation)
and the level of dataset splitting. We then merge some hydrographic basin to end
up with four region of interpolation: West, East, South and North. (XXX ADD 
basins merged)

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their prediction performance. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014, @hoef_ssn_2014].

### Performance and post check

```{r eval = TRUE}
myload(cv_press_interp_mv_avg,
  dir = mypath("data-raw", "polluants"))
myload(cv_temp_press_interp_mv_avg,
  dir = mypath("data-raw", "naiades_temperatures"))
myload(cv_flow_press_interp_mv_avg,
  dir = mypath("data-raw", "flow"))
dbo_cv <- cv_press_interp_mv_avg %>%
  filter(parameter == "dbo5")

cv_dbo_temp_flow <- map2_dfr(
  list(dbo = dbo_cv, temp = cv_temp_press_interp_mv_avg, flow = cv_flow_press_interp_mv_avg),
  c("BOD", "Temperature", "Flow"),
  function(x, y) {
    x %>%
      select(basin, year, bias, std.bias, RMSPE, std.MSPE) %>%
      group_by(basin) %>%
      summarise_if(is.numeric, list(avg = ~mean(., na.rm = TRUE), sd = ~sd(., na.rm = TRUE))) %>%
      mutate(Variable = y) %>%
      select(Variable, basin, everything()) %>%
      ungroup
  })

cv_dbo_temp_flow %>%
  mutate_if(is.double, ~signif(., 2)) %>%
  kable(
    align = "c",
    format = "latex",
    booktabs = T,
    caption = paste0("Results of the cross-validation for the interpolation of
      BOD, temperature and flow. Average over 4 interpolations are shown.") 
      ) %>%
    collapse_rows(columns = c(1, 2), valign = "top") %>%
    kable_styling(latex_options =c("striped", "scale_down"))
```

```{r}
cv_bias_caption <- paste0(
  "Bias of the interpolation from Leave-One-Out Cross Validation (LOOCV). "
)
```

The performance of the interpolation was evaluated with leave one out
cross-validation (already implemented in the SSN package). The value of the
interpolated variable of each station was predicted after removing it from the
dataset.

When predicted values were below the minimum value or above the maximum value
recorded in the dataset for flow, temperature and BOD, we replaced the
predicted value by respectively the minimum and the maximum value found in the
dataset. We ran the interpolation four times and we kept the predicted value
which had the least standard deviation.

# Species list and food diet data

```{r sp-list}
myload(species, fish_diet_shift, dir = mypath("data"))
kable(species,
  format = "latex",
  booktabs = T,
  label = paste0("fish-sp"),
  caption = "Fish species list."
  ) %>%
kable_styling(latex_options =c("striped"))
```

```{r food}
kable(select(fish_diet_shift, - species_name, -remark),
  format = "latex",
  booktabs = T,
  label = paste0("onto-diet"),
  caption = "Ontogenic diet shift for each species"
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

# Analysis

## Time series classication

The CV of biomass over can cover multiple dynamics, including increases,
stable, decreases and a mix of them. Theoretically, however, the CV computed to
quantify stability is often represented with the assumption that the overall
dynamic of the biomass varies around an equilibrium. In order to approach this
condition, we kept only the stations that have met this criteria. To qualify the
biomass trajectories over time, we used the symbolic aggregated approximation
method (SAX) which is commonly used in time series analysis and implemented in
the `jmotif` R package. This method classifies time series rather than
quantifies trends and so is generic enough to encompass the diversity of time
series shapes. The classification necessitates three steps. First, the time
series are normalized (i.e. Substituting the mean and divide by the standard
deviation). The second step is to reduce the dimensionality of the time series
by Piecewise Aggregation Approximation (PAA) method. Each time series is divided
in $N$ frames of equal size that allows to the reduce the dimensionality of the
time series. For each frame, we simply kept the mean value of the time series.
The last step is to qualify each frame into string. Because values of
normalized time series typically follows a normal distribution ($N(0,1)$), a
breakpoints can be computed by dividing the area under the normal distribution
in equal parts. The algorithm attributes a alphabetic letter to each part from
the lowest to the highest ($a$, $b$, $c$, ...). We set three breakpoints, so
$a$, $b$ and $c$. Finally, the trajectory of each time series is classified as
a combination of $a$, $b$, $c$ letters. $abc$ and $cba$ time series represent
respectively monotonic increasing and decreasing trajectories. For data analysis,
we kept only the $bbb$ which represent time series that showed no temporal
tendencies, with respect to our methodology.

## PCA of environmental variables

```{r eig-plot}
rotated_eigen_values <- 
  tibble(
    name = paste0("RC", seq(1,14)),
    type = c(rep("Kept", 2), rep("No kept", 14 - 2)),
    value = pca_rotated$rotated$values
  )
eigen_rotated_plot <-
  ggplot(rotated_eigen_values,
    aes(x = reorder(name, -value) , y = value)) +
  geom_col(aes(fill = type)) +
  scale_fill_discrete(name = "Status") +
  labs(x = "Rotated component", y = "Eigen values")
```

```{r contrib-pca}
#https://stats.stackexchange.com/a/143949
caption_contrib_pca <- paste0(
  "Loadings of variable to rotated components of the PCA.
  Loadings are the coefficients which determine the components of the PCA as
  linear combinations of the variables.
  "
)

loadings_pca <- matrix(
  as.matrix(round(pca_rotated$rotated$Structure, 2)),
  nrow = 14,
  dimnames = dimnames(pca_rotated$rotated$Structure)
  ) %>%
  as_tibble %>%
  mutate(
    Variable = str_replace_all(
      dimnames(pca_rotated$rotated$Structure)[[1]],
      get_pca_var_name_replacement())
  ) %>%
  select(Variable, RC1, RC2)
kable(
  x = loadings_pca,
  format = "latex",
  booktabs = T,
  label = "contrib-pca",
  caption = caption_contrib_pca
  ) %>%
kable_styling(latex_options = c("striped"))
```

```{r}
corr_pca_caption <- paste0(
  "Spearman correlation between environmental variables used in the PCA.
  The variables have been scaled before to compute the correlation."
)

pca_var_correlation <- pca_pressure %>%
  select(-station) %>%
  #mutate_all(scale) %>%
  as.matrix() %>%
  cor(., method = "spearman")

pca_var_correlation %<>% round(., 2)
dimnames(pca_var_correlation) %<>%
  map(., ~str_replace_all(.x, get_pca_var_name_replacement()))

kable(
  x = pca_var_correlation,
  format = "latex",
  booktabs = T,
  label = "corr-pca",
  caption = corr_pca_caption
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r}
prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))

tot_var_expl_pca <- round(sum(pca_rotated$rotated$Vaccounted["Proportion Var", ]) * 100)

pca_caption <- paste0(
  "Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  "(A) Axis 1 was positively related to the river size, axis 2 was related to
  average temperature and negatively related to the river slope and altitude.
  (B) Axis 3 was negatively related to the CV of stream size.
  Axis 4 was related to the CV of enrichment.
  Finally, axis 5 was related to the CV of flow and average enrichment.")
```

```{r pca-rotated, fig.cap = pca_caption}
p1 <- my_pca_plot(xaxis = "RC1", yaxis = "RC2", ctb_thld = .4)
#p2 <- my_pca_plot(xaxis = "RC3", yaxis = "RC4", ctb_thld = .4)
#p3 <- my_pca_plot(xaxis = "RC4", yaxis = "RC5", ctb_thld = .4)

pca_plot <- plot_grid(p1, eigen_rotated_plot, nrow = 1)
save_plot(
  filename = mypath("manuscript", "bef_stability", "figs", "pca_plot.pdf"),
  pca_plot,
  nrow = 1, ncol = 2,
  base_height = 6/2.54, base_asp = 1.6, base_width = NULL 
  )
pca_plot
```

```{r}
print_prop_var_pca <- prop_var_expl_pca[order(names(prop_var_expl_pca))]
```

The five axis of the rotated PCA explained respectively
`r paste0(prop_var_expl_pca, "%", collapse = ", ")` of the variance.
Please see the main text for the interpretation of the axis.

```{r }
rc2 <- tibble(
  station = pca_pressure$station,
  alt = pca_pressure$alt,
  temperature_med = pca_pressure$temperature_med,
  value = pca_rotated$rotated$scores[,"RC2"]
)

rc2_sf <- station_analysis %>%
  select(station = id) %>%
  filter(station %in% rc2$station) %>%
  left_join(rc2, by = "station")

caption_rc2_map <- paste0(
  "Map of the station values for the second axis of the rotated PCA."
)
```

```{r rc2-map, fig.cap = caption_rc2_map}
p +
  geom_sf(data = rc2_sf, aes(color = log(value + abs(min(value)) + 1 ))) +
  viridis::scale_color_viridis() +
  labs(title = "RC2 (Log)") +
  theme(legend.title = element_blank())
```

```{r pca-two-axis, eval = FALSE}
two_axis <- compute_rotated_pca(.data = pca_pressure, naxis = 2)
two_axis$rotated$Vaccounted

cor(two_axis$rotated$score[,c("RC1", "RC2")],
  pca_rotated$rotated$score[,c("RC1", "RC2")])
```

## Model summary and diagnostics

### Data transformation

Additionally, we logged the axis one, two and three to normalize their
distribution and we reversed the third axis to get increasing average
temperature with increasing values of the axis and so decreasing values of
altitude.

```{r}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 5)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 5), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c(rep("Log base 10", 3), rep("No", 2), "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  interpretation = c(rep("", 11)) 
) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "Variable description and transformation"
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

### SEM hypothesis

We linked environmental variables both to community structure and stability
components. Rooted in assembly theory, environment filters species that can
establish in a community and thus affect community structure. (+ add specific
hypothesis about physical and climate). (1) We hypothesized that enrichment
(characterized by DBO) and mean annual temperature will increase species turnover
[@friedpetersen_drivers_2020]. (2) We hypothesized environmental forcing can also
decrease biomass stability directly without going by community structure. First,
instability of habitat and environmental factors may in turn drive biomass
instability [@hansen_climate_2013]. Secondly, average environmental conditions
such as enrichment can increase directly CV by increasing species biomass
fluctuations [@tabi_warming_2019]. Increasing temperature was also hypothesized
to decrease stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

---
# Species richness to community attributes 
---
To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. In particular, species richness decreased connectance
following results from @thebault_stability_2010; @winemiller_must_1989. But
theoretically, some ecological networks can show the inverse relationship
[@winemiller_must_1989]. (6) We hypothesized also that species richness will
modulate average weighted trophic level because of food web assembly processes.
A community with more fish species can support higher trophic level. Finally, we
set up causal relationships from community attributes to stability components.

---
#  Community attributes to stability
---
We hypothesized that (7) species richness decreased synchrony between species
[@loreau_species_2008] and increased $\overline{CV_{sp}}$
[@tilman_biodiversity_2006]. (8a) We expected that average trophic level had a
positive or negative relationship with synchrony, bottom-up effects should
result in positive whereas top-down in negative relationships. We expected a (8b)
negative relationship between average trophic level and $\overline{CV_{sp}}$ if
predators stabilised prey fluctuations or if the increasing part of
biomass in the predators decreased $\overline{CV_{sp}}$ because of their higher
body mass (if mean increase faster than variance). (9a) Connectance was expected
to decrease or increase $\overline{CV_{sp}}$ and (9b) synchrony because of top
down effects or bottom up effects.

## Model linear Relationship

### Variables

```{r}
hist_var_caption <- paste0(
  "Distribution of the variables used in the Structural Equation Modeling. The
  biomass stability, the CV of species (CVsp), PCA1, PCA2, species richness,
  synchrony and total biomass were logged."
)
```

```{r hist-sem-data, fig.cap = hist_var_caption}
sem_data %>%
  select(log_bm_std, log_rich_tot_std, log_sync, log_stab_std, log_cv_sp, log_RC1, log_RC2, t_lvl, ct) %>%
  gather(variable, value) %>%
  mutate(variable = str_replace_all(variable, get_sem_var_name_replacement())) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free")
```

### Colinearity tables

```{r}
colin_sem_cap <- paste0(
  "Variance Inflation Factor for each individual linear model of the structural
  equation models."
) 
```

```{r s-colin-sem, fig.cap = ""}
mod_stab_list <-
  list(
    nlme::lme(log_rich_tot_std ~ log_RC1 + log_RC2,  random = ~ 1 | basin, data = sem_data),
    nlme::lme(ct ~ log_RC1 + log_RC2 + log_rich_tot_std, random = ~ 1 | basin, data = sem_data),
    nlme::lme(t_lvl ~ log_RC1 + log_RC2 + log_rich_tot_std, random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_sync ~ log_rich_tot_std + ct + t_lvl + log_RC1 + log_RC2,
      random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_cv_sp ~ log_rich_tot_std  + ct + t_lvl
      + log_RC1 + log_RC2, random = ~ 1 | basin, data = sem_data),
    lm(log_stab_std ~ log_cv_sp + log_sync, data = sem_data),
    nlme::lme(log_bm_std ~ log_rich_tot_std  + ct + t_lvl
      + log_RC1 + log_RC2, random = ~ 1 | basin, data = sem_data)
  )
names(mod_stab_list) <- c(
  "Species richness",
  "Connectance",
  "Avg trophic level",
  "Synchrony",
  "CVsp",
  "Biomass stability",
  "Total biomass"
)

colin_stab_sem <- map(mod_stab_list, car::vif)

tmp <- map(colin_stab_sem, enframe)

# colinearity table
colin_stab_sem <- tibble(vif = tmp, mod = names(mod_stab_list)) %>%
  unnest(vif) %>%
  rename(`Response variable` = mod) %>%
  mutate(
    value = round(value, 1),
    name = str_replace_all(name, get_sem_var_name_replacement())
    ) %>%
  spread(name, value)

kable(colin_stab_sem,
  format = "latex",
  booktabs = T,
  caption = "Variance Inflation Factor for the linear models used in structural
  equation modeling"
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r}
caption_mod_res <- paste0(
  "Residuals versus fitted values of the linear models used in the structural
  equation models."
) 
```

```{r mod-res, fig.height = 8, fig.cap = caption_mod_res}
library(nlme)
plot_residuals <- function (x, y = "Diagnostic") {
  plot.lme(x, main = y, xlab = "Fitted values", ylab = "Residuals")
}
p_resid_stab_sem <- map2(mod_stab_list, names(mod_stab_list) , plot_residuals)
plot_grid(plotlist = p_resid_stab_sem, ncol = 2)
```

### Biomass and stability models

```{r}
format_coef_table <- function (sem = NULL) {

  params <- sem$coefficients
  colnames(params)[length(colnames(params))] <- "Stars"

  replace_wo_n <- str_replace_all(get_sem_var_name_replacement(), "\n", "")
  names(replace_wo_n) <- names(get_sem_var_name_replacement())

  params %<>%
    mutate_at(c("Response", "Predictor"),
      ~str_replace_all(., replace_wo_n))

  params
}
```

```{r}
est_stab_sem_caption <- paste0(
  "Estimate of the effects for the temporal stability of annual biomass of the structural equation
  model."

)

kable(format_coef_table(stab_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_stab_sem_caption
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r}
est_bm_sem_caption <- paste0(
  "Estimate of the effects for the structural equation
  model of the median of annual total biomass."
)

kable(format_coef_table(bm_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_bm_sem_caption
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r rsq-model}
rm_n <- str_remove_all(get_sem_var_name_replacement(), pattern = "\n")
names(rm_n) <- names(get_sem_var_name_replacement()) 
rsq_table <- map_dfr(mod_stab_list, piecewiseSEM::rsquared, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2)) %>%
  mutate(Response = str_replace_all(Response, rm_n)) %>%
  dplyr::select(Response, Marginal, Conditional, R.squared)
rsq_table_caption <- paste0(
  "R squared of the linear models used in the structural equation modeling."
)
kable(format_coef_table(bm_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_bm_sem_caption
  ) %>%
kable_styling(latex_options = c("striped"))
```

```{r tab-coeff-bm, fig.cap = "Coefficient"}
# Replacement rules for term
rp_est <- c(
  "(Intercept)" =     "intercept",
  "sd\\_\\(intercept\\)\\.basin" = "sd intercept basin",
  "log_rich" =     "Median richness (log base 10)",
  "troph_group3" =     "high trophic group",
  "log_rich:troph_group3" =     "Median richness (log10) : high trophic group",
  "sig01.basin" =     "sd intercept basin",
  "sd_Observation.Residual" =     "residuals",
  "sigma" = "residuals"
)
# Replacement rules for group
rp_group <- c(
  "fixed" = "Fixed",
  "basin" = "Random"
)
```

```{r}
mask_drop_lm <- names(mod_stab_list) %in% "Biomass stability"
random_coeff_sem <- map(mod_stab_list[!mask_drop_lm], random.effects)
test <- getVarCov(mod_stab_list[!mask_drop_lm][[2]])

get_intercept_variance <- function (mod) {
  vc <- VarCorr(mod)
  suppressWarnings(storage.mode(vc) <- "numeric")
  enframe(vc[1,])
}
random_coeff_sem <- map(mod_stab_list[!mask_drop_lm],
  get_intercept_variance) %>%
  enframe(name = "variable") %>%
  unnest(value) %>%
  pivot_wider(names_from = "name", values_from = "value")
get_intercept <- function (mod) {
  mod[["coefficients"]]$fixed["(Intercept)"]
}
intercept_sem <- map_dbl(mod_stab_list[!mask_drop_lm],
  get_intercept) %>%
enframe(name = "variable", value = "Intercept")

formated_random_effects <- intercept_sem %>% 
  left_join(random_coeff_sem, by = "variable") %>%
  rename(`Response variable` = variable)
#format_random <- function (x, y) {
  #tibble(
    #Variable = y, 
    #`Hydrographic basin` = rownames(x),
    #Intercept = x[[1]]
  #)
#}
#formated_random_effects <- do.call(rbind, map2(random_coeff_sem, names(random_coeff_sem), format_random))

caption_random_effect <- paste0(
  "Estimated random effect of hydrographic basin on the intercept of the linear
  models performed in structural equation modeling."
)

kable(formated_random_effects,
  format = "latex",
  booktabs = T,
  caption = caption_random_effect
  ) %>%
kable_styling(latex_options = c("striped"))
```

### Robustness to different dataset selection

```{r sensitivity, message = FALSE, fig.dim = c(10, 7), fig.cap = "Sensivity analysis with different dataset", cache = FALSE}
library(furrr)

myload(op_analysis, op_analysis_wo_holes, op_analysis_ov, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)

myload(network_metrics, dir = mypath("data", "classes"))
sem_sensivity_data <- list(full = op_analysis, bbb = op_analysis_bbb,
  without_holes = op_analysis_wo_holes, bbb_without_holes = op_analysis_wo_holes_bbb)

sem_stab_l_data <- 
  furrr::future_map(sem_sensivity_data,

  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_stab_sem_rich
  )
  )

nb_station <- map_int(sem_sensivity_data, ~length(unique(.x$station)))

sem_stab_results <- map(sem_stab_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_stab_results) <- str_to_lower(colnames(sem_stab_results))
mysave(sem_stab_results, dir = mypath("data"), overwrite = TRUE)
```

```{r sem-sens-bm, cache = TRUE, fig.cap = "Sensivity analysis for Biomass"}
sem_bm_l_data <- furrr::future_map(sem_sensivity_data,
  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_prod_sem_rich)
)
sem_bm_results <- map(sem_bm_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_bm_results) <- str_to_lower(colnames(sem_bm_results))
mysave(sem_bm_results, dir = mypath("data"), overwrite = TRUE)
```

```{r}
sens_cap <- paste0(
  "Sensivity analysis of the Structural Equation models for total biomass
  and its temporal stability."," The Constant dataset is the one kepting only the
  sites that had not a significant temporal trend in biomass; ", "Constant/
    without holes dataset kept only the constant sites plus the ones that had not
  holes in their sampling period; ", "Complete is the dataset containing all the
  site that have been followed ten years or more, as the other dataset;", "Without holes dataset is
  the dataset containing only the sites that had no holes in the their sampling period."
)
```

```{r, fig.height = 8, fig.cap = sens_cap}
p_sens_bm <- plot_sensivity_analysis(.data = sem_bm_results) +
  theme(legend.position = "none")

p_sens_stab <- plot_sensivity_analysis(.data = sem_stab_results) +
  guides(color = guide_legend(direction = "horizontal", nrow = 2))
legend_b <- get_legend(p_sens_stab)
p_sens_stab %<>% + theme(legend.position = "none")
p <- plot_grid(
  p_sens_bm, p_sens_stab, legend_b, nrow = 3,
  rel_heights = c(1, 1, .1),
  labels = c("Total biomass", "Biomass stability")
)
p
```

# References