### 2020-11-17 manuscript/bef_stability/table_onto_species.Rmd

---
title: List of species and ontogenic diet data
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
    toc: false 
fontsize: 12pt
header-includes:
   -  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{pdfpages}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

```{r, echo = FALSE, message = FALSE, collapse = FALSE}

mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

#load(file = mypath("manuscript", "bef_stability", "result", "workspace.rda"))

knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
library(nlme)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))
source(mypath("R", "cleaning_methods.R"))

theme_set(theme_alain())
```

\beginsupplement

```{r sp-list}
myload(species, metaweb_analysis, fish_diet_shift, dir = mypath("data"))
species %<>%
  filter(code %in% metaweb_analysis$species) %>%
  select(-max_size) %>%
  rename(`Common name` = common_name, `Latin name` = latin_name)
kable(species,
  format = "latex",
  longtable = T,
  booktabs = T,
  label = paste0("fish-sp"),
  caption = "Fish species list.") %>%
kable_styling(latex_options =c("striped", latex_options = c("repeat_header")))
```