### 2020-12-17 manuscript/bef_stability/manuscript.Rmd
Decrease to 5000 words:
8391 - 21 - 237 - 2730 - 251 - 151

---
title: Species richness and network structure jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

```{r, echo = FALSE, results = FALSE, message = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE 
)

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))

theme_set(theme_alain())
```

```{r, results = FALSE}
knitr::opts_chunk$set(
  results = TRUE 
)
```

- Author affiliation:
  - Alain Danet: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France
  - Maud Mouchet: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France
  - Willem Bonnaffé: Ecological and Evolutionary Dynamics Lab. Zoology Research
    and Administration Building, 11a Mansfield Rd, Oxford OX1 3SZ, United
    Kingdom.
  - Elisa Thébault: Sorbonne Université, CNRS, IRD, INRAE, Université Paris Est
    Créteil, Institute of Ecology and Environmental Sciences of Paris
    (iEES-Paris), Paris, France  
  - Colin Fontaine: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France

- Author contributions: AD, CF, MM and ET designed the study. WB designed the
  network inference method and collected the food diet data. AD performed the
  data analysis and wrote the first draft of the manuscript. All authors
  contributed substantially to revisions.

- Data accessibility statement: the ONEMA database is available upon request to 
  eddy.cosson@afbiodiversite.fr and thierry.point@afbiodiversite.fr. The code
  used for the analysis and to generate the manuscript is available on github:
  (alaindanet/fishcom.git)

- Article type: Letters

- Running title: Foodweb structure, biomass and stability

- Number of words:
  - Abstract: 151
  - Main text: 4980

- Number of references: 92
- Number of figures: 2
- Number of tables: 1
- Number of text boxes: 0

- Coresponding author:
  - Alain Danet
  - UMR 7204 CESCO, CP 135; 43, rue Buffon, 75005 Paris, France
  - alain.danet@mnhn.fr 
  - +33140798134

# Abstract

Biodiversity-ecosystem functioning (BEF) and foodweb complexity-stability
relationships are central to ecology. However, they remain largely untested in a
foodweb context in natural settings. Here we study the links among
environmental conditions, richness, foodweb structure, annual biomass and its
temporal stability using a standardized monitoring of 99 stream fish communities
spanning from 1995 to 2018. We first reveal that both richness and average
trophic level are positively related to annual biomass, with effects of similar
strength. Second, we show that network structure affects community stability,
with a positive effect of mean trophic level, and contrary to general
expectation, we find that richness destabilizes fish communities. Finally, we
find that environmental conditions affect both biomass and its stability via
effects on richness and network structure. Strikingly, effects on stability are
mediated by changes in species stability rather than synchrony, which contrasts
with results from single-trophic communities and might be characteristic of
multi-trophic communities.

# Introduction

---
#### Biodiversity and ecosystem properties 
---

Current biodiversity losses and further global change have the potential to
affect ecosystems in complex ways [@loreau_biodiversity_2001;
@tilman_biodiversity_2014]. Biodiversity has indeed a tremendous effect on
ecosystem functions and their stability in response to perturbations
[@hector_plant_1999; @loreau_biodiversity_2001; @hooper_effects_2005;
@dunne_network_2006; @duffy_why_2009], because it largely determines both the
amount and the distribution of matter and energy fluxes within and among
ecosystems [@loreau_meta-ecosystems:_2003]. Understanding the relationship
between biodiversity and ecosystem properties is thus crucial to better
anticipate the impacts of global change.

---
#### Species richness increases total biomass
---

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006].  Species richness has
been shown to promote higher biomass production and  higher temporal stability
of total biomass in a variety of ecosystems and taxa.  Indeed species richness
generally increases primary productivity and total biomass in experimental plant
communities [@tilman_productivity_1996; @grace_integrative_2016;
@hector_plant_1999; @tilman_biodiversity_2014; @grace_integrative_2016], but
also in briophyte communities [@mulder_physical_2001], stream feeders,
[@cardinale_species_2002] and aquatic bacterial consumers
[@pennekamp_biodiversity_2018].  In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020].  Despite the diversity of
ecosystems and taxa studied, the relationship between diversity and ecosystem
functioning in experimental and empirical studies remains mostly focused on
single trophic levels, ignoring the effects of foodweb structure.  This gap is
even more striking for the relationship between diversity and temporal stability
of ecosystem functioning, in spite of long-standing interest for the links
between foodweb complexity and stability in theoretical ecology [e.g.
@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and foodweb structure should modulate ecosystem properties such as
total biomass and its stability. Complementarity in the use of
resources, a main mechanism previously highlighted in single-trophic studies, is
important to understand the effects of species richness in foodwebs [@thebault_food-web_2003;
@poisot_trophic_2013; @schneider_animal_2016], but in that case it also depends
on foodweb connectance [@thebault_food-web_2003; @poisot_trophic_2013].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in species temporal variations is also explaining a significant part
of the stabilizing effects of species richness on total biomass in foodwebs
[@thebault_trophic_2005]. Indeed, a low synchrony in species fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a species is compensated by the increase of an other. However,
higher connectance can reverse the positive effects of species
richness on temporal stability of total biomass, by leading to a negative
relationship between species richness and species temporal variability
[@thebault_trophic_2005]. In addition, the presence of species at higher trophic
levels increases total primary production [@wang_biodiversity_2018]. The
addition of predators has also cascading effects on both species variability and
synchrony at lower trophic levels [@teng_dynamics_2004;
@shanafelt_stability_2018]. Overall, these theoretical studies show that the
effects of species richness and network structure, such as connectance and
average trophic level, are potentially conflicting or synergistic in such a way
that they need to be addressed jointly.

The gap in studies that consider jointly effects of diversity and foodweb
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and foodweb complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that describing
trophic interactions between species is highly labour intensive. These
challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires time-series data and longer-term
analyses. In this context, recent developments to predict species trophic
interactions [@gravel_inferring_2013], combined with observational long-term surveys
of ecological communities, can offer an opportunity to investigate the links
between diversity, foodweb structure and the stability of total community
biomass in natural settings.

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
foodweb structure. For instance, larger ecosystems host foodwebs with higher
diversity and trophic levels [@mchugh_dual_2010] while temperature can affect
foodweb diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. When analyzing
data from natural communities, the relationships among species richness, total
biomass and its temporal stability might also arise from effects of the
environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness.

---
#### Stream fish communities is a good model to study effects of horizontal & vertical diversity
---

Because their trophic interactions are well known, stream fish communities
constitute a good model to study the effects of network structure on total
biomass and its temporal stability. Despite that fish trophic interactions are
difficult to measure in natural conditions, a substantial body of literature
supports that they can be inferred through body size and diet informations
[@jennings_weak_2001; @gravel_inferring_2013], especially
in stream communities including fishes [@brose_predator_2019]. Lastly, most of
the diversity-biomass and diversity-stability relationships, have been assessed
in terrestrial ecosystems while there is a evident need to test such
relationships in a variety of ecosystem types.

Our aim is to document the relationships among species richness, foodweb
structure, total biomass and its temporal stability in natural communities. To
do so, we used the ONEMA dataset consisting in a standardized sampling of fish
communities in French streams [@edeline_ecological_2013]. We selected 99 sites
that have been sampled for more than 10 years spanning on the 1995-2018 period. We
inferred 1496 interaction networks, based on body size and diet. We question
whether species richness, connectance and average trophic level are related or
not to total biomass and its temporal stability. We further question whether
species richnes and foodweb structure have antagonistic or mutual effect on
total biomass and its temporal stability. Finally, we question how ecosystem size
and average temperature are related to total biomass and its temporal stability.

# Methods

```{r, cache = FALSE}
myload(qt_station, dir = mypath("data"))
matmet_cap <- paste0( 
  "(A) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. The red dots show the location of (C) and (D) sites.
  ", " Inner borders draw the limits of the hydrographic basins. ", "(B) The
  metaweb of the study. Each trophic species (N = 412) is represented by a node
  and each
  color represents a species (S = 45). Resource nodes (7) are drawn in grey. ",
  "The species names corresponding to the three digit codes are found in Table
  S1." ,
  "Yearly temporal series of networks of two stations having respectively a low
  (C) and high (D) temporal stability of biomass. C and D stations have
  respectively biomass stability the closest to the first and third quartile of
  the distribution of the temporal stability of biomass in the dataset,
  respectively ", round(qt_station$bm_std_stab[1],2), " and ",
  round(qt_station$bm_std_stab[2],
    2), ". Point size is proportional to the biomass of the trophic species.")

```

```{r}
# Basin
st_basin <- get_basin_station(sf_obj = FALSE)

# Select fishing operation 
myload(op_analysis, op_analysis_wo_holes, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
range_surface <- map_dbl(c(min, max), ~.x(op_analysis_bbb$surface, na.rm =TRUE))
range_surface[1] <- ceiling(range_surface[1] / 10) * 10
range_surface[2] <- ceiling(range_surface[2] / 100) * 100
# quantile(op_analysis_bbb$surface, na.rm =TRUE)
```

## Data preparation and filtering

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). The stations where the stream
was of intermediate size could be sampled alternatively with one of the two
sampling protocols. In this case, we kept only the most frequently used sampling
protocol. The sampling season spanned from late spring to autumn. There was low
heterogeneity of the sampling month within stations, with the median sampled
month being mid-August while the median of the standard deviation of the sampled
month within station was $0.7$.

We selected the stations that had been sampled ten years or more. As the concept
of community variability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only
stations with no temporal trends in total biomass were selected (detailed in
appendix B, part 2). However, the results were robust to the relaxation of this
hypothesis (Fig. S7). These selection steps resulted in 99 stations distributed
over seven hydrographic basins (Fig. \@ref(fig:matmet) A).

## Total biomass and its temporal stability

### Biomass estimation

The body size of all fish individuals was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Species biomass and total biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, species biomass and
total biomass were corrected by the surface sampled (range `r range_surface[1]`-
`r range_surface[2]` square metres), and expressed in grams by square meter
($gm^{-2}$). In the analysis, the total biomass of a site has been defined as
the median of the total biomass across sampling years.

### Stability measures

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the biomass over
time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being
respectively the standard deviation and the temporal average of the annual total
biomass of the station $i$). Then, temporal stability was equal to $CV^{-1} =
\mu \sigma^{-1}$.

We described the mechanisms driving the temporal stability of biomass, by
decomposing the $CV$ of total biomass ($CV_{com}$) into a synchrony ($\phi$) and
a weighted average species $CV$ ($\overline{CV_{sp}}$) components according to
@thibaut_understanding_2013:

$$
CV_{com, i} = \overline{CV_{sp, i}} \times \sqrt{\phi_i}
$$

with $CV_{com, i}$ being the Coefficient of Variation of the biomass in the
station $i$, $\overline{CV_{sp, i}}$ the average coefficient of variation of
species in the station $i$ weighted by the relative biomass of the species, $B_{k,i}$.
Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
total biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_k}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

$$
\overline{CV_{sp, i}} = \sum_k{\frac{B_{k,i} {CV}_{k, i}}{\sum_k{B_{k,i}}}}
$$

$$
\phi_i = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sqrt{\sigma^2_{x_{k,i}}}})^2} = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sigma_{x_{k,i}}})^2}
$$

## Species richness and foodweb structure

### Species richness

Species richness in a given station has been defined as the total number of species present across
all the sampling events. We controlled species richness for sampling effort by
dividing the total number of species by the total surface sampled in the given
station. Species richness was then expressed in number of species per square
meter sampled.

### Network inference

```{r}
myload(metaweb_analysis, dir= mypath("data"))
myload(trophic_class, dir= mypath("data", "classes"))
meta <- metaweb_analysis
#metaweb_analysis$resource
#meta$metaweb[meta$resource, meta$resource]
```

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Fig.
\@ref(fig:matmet) B), then (2) determine network of a given community by
extracting the metaweb according to the trophic species present at a given
site and sampling event (Fig. \@ref(fig:matmet) C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, alvins feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet) B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. The ontogenic fish diet
database was filled according to fishbase and literature (appendix A, Table S2). Each
fish species had two or three life stages which are delimited by a body size
range (Fig. S2). The life stage of a fish trophic species was determined with
the center of its body size range, hereafter midpoint (Fig. S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic specie was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint was included in the
predation window (Fig. S2). The fish-resource trophic links were set between
the trophic species and the resources nodes according the food diet of their
size class, for both the piscivorous and non piscivorous trophic species.
Lastly, the trophic links among resource nodes were set according to the
literature [@allan_stream_2007; @hart_handbook_2008].

### Network structure

For each trophic network, corresponding to a particular site and year, we
computed two of the most commonly used network metrics to characterize food
webs, namely the connectance and mean trophic level. Connectance describes the
level of generalism in the network and is calculated as the actual number of
interactions divided by the maximum number of interactions if all species interact
together ($C = L / N^2$, L and N being resp. the number of links and of trophic
species plus resource nodes). The mean trophic level $\overline{T}$ was computed
as the mean trophic level weighted by the biomass of each trophic species
(excluding resource nodes for which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times (\sum_i^N{B_i})^{-1}}$, with $T_i$ and
$B_i$ being respectively the trophic level and the biomass of the node $i$. The
trophic level of the nodes was computed as one added to the average trophic
level of its food item. For each site, the median of connectance and average
trophic level over time was used in subsequent analysis.

```{r, fig.show = "hide"}
myload(
  hab_analysis,
  temporal_station_desc,
  temporal_press_polluants,
  geo_station,
  dir = mypath("data")
)
habitat_press <- build_habitat_pressure_dataset(
  .habitat_analysis = hab_analysis,
  .tmp_press = temporal_press_polluants,
  .tmp_st_desc = temporal_station_desc,
  .geo_st = geo_station 
)

# Make PCA 
habitat_press %<>% na.omit()
mask <- names(habitat_press) %in% names(get_pca_var_name_replacement())

pca_rotated <- compute_rotated_pca(
  .data = habitat_press[, mask],
  naxis = 2
)
habitat_pressure <- 
  cbind(habitat_press[, c("station", "temperature_med", "alt")], pca_rotated$rotated$score) %>%
  as_tibble
ggplot(habitat_pressure, aes(x = RC2, y = temperature_med)) +
  geom_point()

ggplot(
  pivot_longer(habitat_pressure, cols = c("RC1", "RC2"),
    names_to = "variable", values_to = "value"
  )
  , aes(x = value)) +
geom_histogram() +
facet_grid(cols = vars(variable))

mysave(habitat_press, habitat_pressure, dir = mypath("data"), overwrite = TRUE)

```

```{r}
prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))
tot_var_expl_pca <- sum(prop_var_expl_pca) 
```

```{r, message = FALSE, results = FALSE}
# Com
com_data <- compute_community_temporal_analysis(.op = op_analysis_bbb,
 type_network_metrics = "classes"
)
# Compute sem dataset
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync_std"]],
  nmds = NULL,
  basin = st_basin
  )
mysave(sem_data, dir = mypath("data"), overwrite = TRUE)
```

## Environmental variables

We characterised the sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow. As data from the naiades database are
not collected on the same site as our fish sampling sites, we performed an
interpolation of the water temperature and flow based on the spatial stream
network [@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure
is detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of water temperature, flow, Biological Oxygen Demand (BOD), river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Fig. S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, DBO and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

## Statistical analysis

Structural Equation Modeling enables to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and indirect, between variables
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. We defined a structural equation model explaining the
average annual total biomass as a function of the two synthetic environmental
variables, species richness and foodweb structure as measured by connectance and
average trophic level [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and network
structure directly affect total biomass (Fig.
\@ref(fig:sem-plot) A)
and that species richness affected foodweb structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure for temporal stability of biomass (Fig.
\@ref(fig:sem-plot) C). Instead of the annual total biomass, we modelled the
synchrony and population variability which in turn determined the temporal
stability of biomass.

The structural equation models were based on Gaussian linear models. All variables
except the connectance and average trophic level
were logged to fulfill model assumptions (see residual plots and data
transformation in Fig. S5 and S6, Table S5). We set the seven
hydrographic basins as a random effect on the intercept to
account for their heterogeneity. We checked for the absence of multicollinearity 
with Variance Inflation Factor (VIF, Table S6). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). The paths whose
slope had a $P$ value inferior to $0.05$ were reported in the main text, all the
path values are presented in Table S7 and S8.

Indirect effects on total biomass and its temporal stability
were computed by multiplying slopes along the path. As an example, if we consider
$A \to B \to C$, the indirect effect of A on C is equal to $r_{AB} \times
r_{BC}$. Only significant direct slopes were entered in the calculation of
indirect effects. Total effects were computed as the sum
of direct and indirect effects.

Environmental variables were interpolated with the `openStars` and `SSN`
R packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the `nlme` and `piecewieseSEM` R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and the manuscript are accessible on github (alaindanet/fishcom). The analyses were
performed with R version 3.6.3 [@r_core_team_r_2020].

```{r get-data}
library(piecewiseSEM)
library(lme4)
library(nlme)

#ctrl <- lmeControl(opt='optim' relative)
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optimx'))
```

```{r}
# Compute SEMs
stab_sem_rich <- compute_stab_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))

bm_sem_rich <- compute_prod_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))

def_env_x_pos <- c(from = 0, to = 3)
asp <- 1.6

library(DiagrammeR)
meta_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability","figs", "meta_stab_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "stab_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)

meta_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "meta_bm_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "bm_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)
```

```{r get-sem-coeff}
coef_stab_sem <- coef(stab_sem_rich) %>%
  select(- 9) %>%
  filter(P.Value < 0.05)

coef_bm_sem <- coef(bm_sem_rich) %>%
  select(- 9) %>%
  filter(P.Value < 0.05)
```

```{r, cache = FALSE}
sem_caption <- paste0(
  "Metamodel of the Structural Equation Models for the (A) median of annual
  total biomass (in grams by square meter) and (C) its temporal stability.",
  "Results of the Structural Equation Models for
  (B) the median of annual total biomass and (D) its temporal stability. Red and black
  arrows represent respectively negative and positive path coefficients. The
  path coefficients are standardized. For convenience, only significant
  paths (i.e. $P < 0.05$) are shown. Rsq: Ajusted $R^2$; Avg:
  Average, CVsp: weighted temporal CV of the population; PCA1 and PCA2: respectively the first and
  second axis of the PCA performed on environmental variables (see methods)."
)
```

```{r get-bm-sem-single-effect}
bm_sem_coef <- filter(coef_bm_sem, P.Value <= 0.05)

rich_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_rich_tot_std")
rich_on_tlvl <- get_clean_sem_coef(sem = bm_sem_coef, resp = "t_lvl", pred = "log_rich_tot_std")
rich_on_ct <- get_clean_sem_coef(sem = bm_sem_coef, resp = "ct", pred = "log_rich_tot_std")
tlvl_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "t_lvl")
ct_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "ct")
RC1_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC1")
RC2_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC2")
```

```{r get-bm-tot-effect-env}
bm_all_params <- get_coefficient_piecewisesem(sem = bm_sem_rich, p_val_thl = NULL)

# effect of env on bm 
env_on_bm <- get_env_on_bm(fit = bm_all_params, p_val_thl = 0.05,
  type = "std")
# rich on bm 
rich_on_bm <- get_rich_on_bm(fit = bm_all_params,
  p_val_thl = 0.05, type = "std")
# direct
ct_on_bm_trick <- list(direct = c(ct = ct_on_bm), total = c(ct = ct_on_bm))
tlvl_on_bm_trick <- list(direct = c(t_lvl = tlvl_on_bm), total = c(t_lvl = tlvl_on_bm))

tmp_bm_indir_table <- map(list(env_on_bm, rich_on_bm, ct_on_bm_trick,
    tlvl_on_bm_trick),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```

```{r}
stab_all_params <- get_coefficient_piecewisesem(sem = stab_sem_rich, p_val_thl = NULL)
# effect of env on stab
env_on_stab <- get_env_on_stab(fit = stab_all_params, p_val_thl = 0.05,
  type = "std")
rich_on_stab <- get_rich_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
ct_on_stab <- get_ct_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
tlvl_on_stab <- get_tlvl_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")

tmp_stab_indir_table <- map(list(env_on_stab, rich_on_stab, ct_on_stab, tlvl_on_stab),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```
```{r get-sem-single-effect}
stab_sem_coef <- filter(coef_stab_sem, P.Value < 0.05)

sync_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_sync")
cv_sp_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_cv_sp")
RC1_on_sync <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_sync", pred = "log_RC1")
tlvl_on_cv_sp <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_cv_sp", pred = "t_lvl")
```

```{r get-sem-tot-effect-env}
# effect of env on sync and cv_sp
env_on_sync <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_sync", type = "std")
env_on_cv_sp <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_cv_sp", type = "std")
# env on stab via network 
env_on_t_lvl <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "t_lvl", type = "std")
env_on_ct <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "ct", type = "std")
env_on_richness <- get_env_on_rich(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
```

# Results

The structural equation models explained a large part of the variation
of species richness, connectance, synchrony, $\overline{CV_{sp}}$ and total
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 =
.11$, Fig. \@ref(fig:sem-plot) B and D). Species richness had a
significant negative effect on mean trophic level ($r_\delta = 
`r rich_on_tlvl`$) and connectance ($r_\delta = `r rich_on_ct`$, Fig.
\@ref(fig:sem-plot) B and D), meaning that a part of the effects of species
richness went through network structure.

## Species richness and average trophic level are positively correlated with total biomass

```{r get-sem-tot-effect-rich}
rich_on_sync <- get_rich_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_sync", type = "std")
rich_on_cv_sp <- get_rich_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_cv_sp", type = "std")
```

Species richness and mean trophic level had a significant positive direct effect
on total biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-plot) B, Table
\@ref(tab:indir-sem)), while connectance had no direct effect on total biomass
(Fig. \@ref(fig:sem-plot) B).

The two PCA axes capturing environmental variables had no significant direct
effects on total biomass, but indirect (Table \@ref(tab:indir-sem)). The first
axis, PCA1, related to the stream size was positively linked to total biomass
($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through both species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). The second axis, PCA2, related to the average
annual temperature and inversely to altitude of the sites was positively related
to total biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Table \@ref(tab:indir-sem)). The total effect
of PCA1 and PCA2 on total biomass was of the same magnitude than the total effects of species
richness on total biomass.

## Temporal stability of annual biomass

By its effects on synchrony and $\overline{CV_{sp}}$, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). The negative effect of $\overline{CV_{sp}}$ on temporal
stability was twice higher than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Fig.
\@ref(fig:sem-plot) D).

The total effect of mean trophic level on temporal stability of biomass was of
the same magnitude than the one of species richness (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Fig. \@ref(fig:sem-plot) D) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-plot) D, Table \@ref(tab:indir-sem)).

The total effect of PCA2, related to average annual temperature and inversely to
altitude, on temporal stability was of higher magnitude than the effects of
species richness and mean trophic level (Table
\@ref(tab:indir-sem)).
The PCA2 had a total negative effect on temporal stability, meaning that higher
temperature results in lower stability
($r_\delta = `r env_on_stab$total["log_RC2"]`$, Table \@ref(tab:indir-sem)). 
PCA2 had an indirect positive effect on temporal stability through a
direct positive effect on mean trophic level ($r_\delta = 
`r env_on_stab$via_tlvl["log_RC2"]`$, Table \@ref(tab:indir-sem)). However, PCA2
had a negative effect on temporal stability through a direct positive effect on 
$\overline{CV_{sp}}$ and species richness (resp.  $r_\delta = 
`r env_on_cv_sp$direct["log_RC2"]`$ and $r_\delta = `r env_on_stab$via_richness["log_RC2"]`$,
Table \@ref(tab:indir-sem)). This was also true for the PCA1, related to stream
size, with a resulting very slight positive effect of PCA1 on temporal stability
($r_\delta = `r env_on_stab$total["log_RC1"]`$, Table \@ref(tab:indir-sem)).

```{r, cache = FALSE}
cap_sem <- paste0(
  "Direct, indirect and total effects of environment, species richness,
  connectance and mean trophic level on total biomass
  and its temporal stability. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable.
  Indirect effects are considered as one step from the target variable. For
  exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.62$. Only the
  significant direct effects were retained for the computation of indirect
  effects. ", "As an example, the indirect effect via species richness takes
  into account the direct effect of species richness as well as the effect of
  species richness through connectance and average trophic level. ",
 "NA values indicate that the corresponding paths do not exist in the
 specified Structural Equation Model (SEM). ", "NS: statistically non-significant."
)
```

```{r indir-stab, cache = FALSE}

full_table <- bind_rows(
  tmp_bm_indir_table,
  tmp_stab_indir_table
) %>%
mutate(`SEM type` = c(rep("Total annual biomass", 5), rep("Temporal Stability of annual biomass", 5))) %>%
select(`SEM type`, variable, direct, via_richness, via_ct, via_tlvl, via_cv_sp, via_sync, total) %>%
format_table() %>%
select(`SEM type`, everything())

full_table %<>%
  mutate(Direct = ifelse(Direct == 0, "NS", Direct)) %>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)

library(kableExtra)
mytable <- kable(full_table,
  format = "latex",
  booktabs = T,
  label = paste0("indir-sem"),
  caption = cap_sem
  ) %>%
kable_styling(latex_options =c("scale_down")) %>%
add_header_above(c(" " = 3, "Indirect effects via" = 5)) %>%
column_spec(1:2, width = "3cm") %>%
collapse_rows(columns = 1, valign = "top")
#write(mytable, file = mypath("manuscript", "bef_stability", "table.tex"))
```

# Discussion

---
#This study is unique by its level of integration, from environmental
#effects to total biomass and its stability, going through species richness and
#foodweb structure. It is the first direct assessment of a possible joint effect
#of species richness and network structure on both total biomass and its temporal
#stability. We found that species richness and mean trophic level had a positive
#effect on total biomass. In contrast, species richness had a negative effect on
#temporal stability whereas mean trophic level covaried positively. We also found
#that the effect of species richness and mean trophic level on total biomass and
#its temporal stability were not independent as species richness had a negative
#effect on mean trophic level. Furthermore, we found that connectance had no
#significant effect on both total biomass and its temporal stability. The results
#shows that temporal stability was mediated by changes in species variability 
#rather than synchrony. Finally, we found that the magnitude of the effects of
#average temperature and altitude was equal or superior to the effects of species
#richness and mean trophic level. 
---

## Species richness increased total biomass but decreased its temporal stability

We found that species richness increased total biomass as already found in salt
marshes, grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018] as well as in streams
[@woods_testing_2020]. The positive effect of species richness on total biomass
might be due to resource use complementarity [@carey_determining_2011], which is
held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or to the higher probability of getting high biomass
species with higher species richness [@loreau_biodiversity_2001]. Our result is
also in agreement with a recent study predicting a positive relationship between
animal diversity and total consumer biomass in size-structured foodweb models
[@schneider_animal_2016].

Surprisingly, we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006; @valencia_synchrony_2020], mesocosms
[@pennekamp_biodiversity_2018] with studies on natural communities such as
butterflies  or lotic fishes [@olivier_urbanization_2020; @franssen_annual_2011;
@houlahan_priority_2017], where positive relationships were found between temporal
stability and species richness. However, some other studies also found a
negative relationship, for instance in natural grasslands [@yang_effects_2011]
or in natural forests as a marginal trend [@declerck_species_2006].

The first step to identify the mechanisms behind those contrasting patterns is to
look at the effects of species richness on synchrony and $\overline{CV_{sp}}$.
As most studies on experimental grasslands
[@isbell_biodiversity_2009; @roscher_identifying_2011] and natural communities
of birds, grasslands, bats and butterflies [@bluthgen_land_2016;
@olivier_urbanization_2020], we found that species richness decreases synchrony,
thereby increasing stability [but see @declerck_species_2006;
@valencia_synchrony_2020]. Our study thus extends the evidence that species
richness decreases synchrony to the case of trophic communities in natural
settings. Species richness can reduce synchrony when (i) a given increase in species
richness raises demographic stochasticity by decreasing population
sizes, or when (ii) species have different responses to environmental variation
[@loreau_species_2008].

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
studies focusing on single-trophic communities have reported mixed effects of
species richness on $\overline{CV_{sp}}$, with destabilising effects of species
richness at population level in grasslands [@jiang_different_2009;
@gross_species_2014], a previous meta-analysis suggested that species richness had
often a positive effect on population stability in multi-trophic communities
[@jiang_different_2009].
In foodwebs, theoretical models predict that the relationship between temporal
stability of populations and species richness depends on interaction strength
and its distribution [@mccann_reevaluating_1997; @thebault_trophic_2005;
@downing_temporal_2020], high interaction strength inducing a negative
relationship whereas the presence of weak interactions at high species richness is
expected to promote stability. Therefore, it is possible that our results are
due to a high degree of omnivory in foodwebs which combined with strong trophic
interactions lead to more fluctuations in population biomass.

---
#@thibaut_understanding_2013 proposed that a
#strongly uneven biomass distribution across populations and that the biomass
#variance of population increases much faster than its mean can generate a
#positive effect of species richness on population ($\overline{CV_{sp}}$)
#stability.
---

## Average trophic level increased total biomass and its temporal stability while connectance had no effect

The average trophic level was positively linked to total biomass. Furthermore,
@wang_biodiversity_2018 have shown that total nutrient uptake might increase
with the length of trophic chains and so results in higher total biomass.  While
the predictions of @wang_biodiversity_2018 were restricted to food chains, and
thus might not hold in more complex foodwebs with omnivorous species, our
results are in agreement with their model predictions and highlight the
importance of food chain length as a determinant of total biomass in foodwebs.
In addition, mean trophic level had a positive effect on biomass stability by
its negative effect on $\overline{CV_{sp}}$, thereby increasing average species
stability. It might result from a higher stability of higher trophic level as
predicted by food chain models in the presence of strong top-down control
[@shanafelt_stability_2018; @barbier_pyramids_2019]. An alternative
explanation is that higher mean trophic level leads complexify biomass pathways
through the foodwebs, which might reduce biomass fluctuations in lower trophic
levels [@mccann_diversitystability_2000].

In contrast, connectance had no effect on total biomass. One might have expected
that an increase in connectance may translate in a higher total nutrient uptake
and so results in higher total biomass in the fish compartment. 
However, some studies suggest that a greater connectance should decrease
complementarity in resource use between species, which is expected to
lower total biomass [@thebault_food-web_2003; @poisot_trophic_2013].
A reason for such an absence of an effect might be explained by the fact that
our data lack information about the lower compartments of the community, i.e.
the primary producers, zooplankton and zoobenthos.

Regarding temporal stability, our results
contrast with numerous previous theoretical studies on the relationship between
connectance and stability [e.g., @angelis_stability_1975; @may_will_1972;
@thebault_stability_2010; but see @neutel_reconciling_2007], although they
considered different stability concepts and metrics [@arnoldi_resilience_2016].
The lack of effect of connectance on the temporal stability of total biomass
might be related to the fact that our measure of connectance does not take into
account the distribution of interaction strengths in the foodwebs. Indeed,
foodweb connectance is expected to have different consequences on the temporal
stability of total biomass depending on the strength of trophic interactions
[@thebault_trophic_2005]. However, our measure of connectance does not fully
account for the whole foodweb structure, and further research should examine the
effects of connectance on total biomass and its temporal stability in more
complete empirical foodwebs.

One limitation of our study is indeed the use of inferred networks, i.e. not
observed interactions, and the lack of of information on the other compartments
of the foodweb (e.g. primary producers, decomposers, the benthic part of the
foodweb). However, this approach is a promising way to explore how the
architecture of ecological networks vary along environmental gradients across
time. It constitutes an interesting complement to the real assessment of
ecological interactions. The use of inferred networks also allow to extract most
of the information contained from already available long term datasets.

## Population stability was the main driver of temporal stability

In contrast with previous studies, we found that the average population
variability ($\overline{CV_{sp}}$) had a stronger negative effect on biomass
stability than synchrony [@tilman_biodiversity_2006;
@olivier_urbanization_2020]. It explains why we observe a negative effect of
species richness on temporal stability of biomass. The lower effect of synchrony
might be explained by the presence of generalist predators in stream
communities, that raise synchrony between prey species that share the
same predator and thus reduces compensatory
population dynamics [@raimondo_interspecific_2004]. Alternatively, it is
possible that, in aquatic foodwebs, the occurrence of large oscillatory dynamics
 leads to a higher contribution of population
variability rather than synchrony to community stability
[@mccann_reevaluating_1997]. How both strengths and the relative contribution of
synchrony and population variability to overall stability vary between
foodwebs and single trophic communities is still an open question.

---
#Rough average synchrony in Théophile study: exp(-0.75) = .47 or exp(-0.6) = .54
#This study: sync avg = .69
---

## Environment had an strong effect on total biomass and its temporal stability

We found that the total effect of environmental drivers on total biomass and its
stability was mainly indirect and of the same magnitude order than the effect of
species richness and average trophic level, highlighting that environmental
changes may result in significant ecosystem responses, through alteration of
both species richness and network structure.

We found that environmental drivers overall affect total biomass indirectly
[@woods_testing_2020] rather than directly , contrasting with previous studies
[@yvondurocher_warming_2011; @dossena_warming_2012; @grace_integrative_2016;
@maureaud_biodiversityecosystem_2019]. First, consistent with research showing
that foodwebs in wider streams support more species and biomass
[@allan_stream_2007], we found that the PCA axis related to stream size had a
positive effect on total biomass through a positive effect on mean trophic level
and species richness. Greater ecosystem size might provide a greater diversity
of resources through habitat heterogeneity, and so higher species richness and
total biomass by surface unit, through resource complementarity. Greater
diversity of resources can also maintain more biomass at high trophic level
[@post_ecosystem_2000; @doi_resource_2009]. Second, the positive effect of the
PCA axis related to temperature on total biomass was of the magnitude of the
effects of network structure and species richness, in line with the mounting
body of work reporting important effects of temperature on the structure and
biomass of aquatic systems [e.g.  @dossena_warming_2012;
@edeline_ecological_2013; @emmrich_geographical_2014; @hattab_forecasting_2016;
@woods_testing_2020]. We found that warmer streams
supported higher diversity as well as higher average trophic level and total
biomass, which is in agreement with recent findings on streams where the
positive effects of temperature were related to larger nutrient supply
[@ogorman_unexpected_2017].

While the PCA axis related with stream size had only a very weak effect on the
temporal stability of total biomass, the PCA axis related to temperature had a
negative effect on the temporal stability through a direct destabilizing
effect on populations and through indirect effects. Temperature can have both
stabilizing and destabilizing effects in resource-consumer models depending on
the thermal niche of the species and on how nutrient supply varies with
temperature [@uszko_effects_2017]. In our case, the observed increased
variability of populations might also relate with ecosystem enrichment
[@ogorman_unexpected_2017] because average Biological Oxygen Demand covaries
with temperature in our study. Enrichment can increase CV by
increasing species biomass fluctuations while higher temperature might decrease
 stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

---
#Interestingly, the two PCA axes related to environmental conditions had indirect
#effects on stability through two conflicting pathways, either stabilising, via
#network structure, or destabilising, via species richness. The first pathway
#consists in a positive effect of temperature and stream size on mean
#trophic level, in line with the idea that wider streams and warmer conditions
#are favourable to mesotrophic predators [@allan_stream_2007;
#@emmrich_geographical_2014], which in turn stabilises biomass fluctuations,
#possibly due to increased top-down regulation. The second pathway consists in a
#positive effect of temperature and stream size on species richness, which itself
#has a destabilising effect. This result is interesting in that both
#pathways had not been evidenced to jointly play, to the best of our knowledge.
#Furthermore, it is unclear which pathway is dominating, and under which circumstances,
#opening an interesting avenue for further investigations.
---

---
## Final thoughts about empirical foodwebs

#It was unexpected that the average population $\overline{CV_{sp}}$
#had a stronger negative effect on biomass stability than synchrony
#[@bluthgen_land_2016; @olivier_urbanization_2020], and explain why we observe a
#negative effect of species on temporal stability of biomass. The lower effect of
#synchrony can be explained by the  presence of generalist predators in stream
#communities, which can raise synchrony between prey species which shares the
#same predator [@raimondo_interspecific_2004]. Uneven population biomass within
#communities also raise synchrony between species [@thibaut_understanding_2013].

#As found in other studies, we found that species richness was negatively related
#to mean trophic level and connectance [@winemiller_must_1989;
#@dunne_network_2006; @thebault_stability_2010]. Species richness had thus a
#consistent destabilising effect on temporal stability of biomass directly and
#through average trophic level, but it has a direct positive effect on total annual
#biomass directly but a negative one through the decrease of average trophic
#level. It highlights the importance to consider several aspects of
#network structure to order to improve our understanding of the drivers of
#community structure and stability.
---

# Conclusion

Our work contributes to bridging the gap between theoretical, experimental and
empirical findings on the functioning and stability of communities. Our study
extends the assessment of the effect of species richness on total biomass and
its temporal stability to the case of food-webs in natural settings and
highlights the need to consider species richness, network structure and their
dependence to understand both functioning and stability of communities. Further,
we found that the effects of environmental gradient related to temperature and
ecosystem size are largely mediated by their impact on richness and average
trophic level. Finally, our results highlight species variability as key in
transmitting effects of richness, network structure and environmental gradients
on community stability. This contrast with previous results found in single
trophic communities, mostly plants, that mainly focus on synchrony and, although
further confirmation are needed, might be characteristic of animal and/or
multi-trophic communities.

```{r}
save.image(file = mypath("manuscript", "bef_stability", "result", "workspace.rda"))
```

# References {-}

<div id="refs"></div>

\clearpage

# Figures and Table {-}

```{=latex}
\input{table.tex}
```

```{r matmet, cache = FALSE, fig.cap = matmet_cap, results = "fig.show", out.height="80%"}
library(knitr)
include_graphics(path = mypath("manuscript", "bef_stability", "figs", "fig1.pdf"))
```

```{r sem-plot, out.width="\\maxwidth", fig.cap = sem_caption, results = "fig.show", cache = FALSE}
#out.height="120%",
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple

#plot_grid(
  #meta_bm_sem_rich, p_bm_sem_rich,
  #meta_stab_sem_rich, p_stab_sem_rich,
  #labels = LETTERS[1:4], nrow = 2,
  #rel_widths = c(1, 1)
#)
# Put the panel of Colin
arrow_dir <- mypath("manuscript", "bef_stability","fleches", "figure_sem2")
filename <- list.files(arrow_dir, full.names = TRUE)
mask <- str_match(filename, "\\.svg")[, 1] %>% is.na %>% !.
#filename[mask]
type <- c("bm", "meta_bm", "meta_stab", "stab")
sem_list <- map(filename[mask], function (x) {
  w <- 2400
  img <- magick::image_read_svg(
    path = x, width = w, height=w*1.618)#, density = 600
  p_sem <- cowplot::ggdraw() +
    cowplot::draw_image(img)#, scale = 1.0
  return(p_sem)
})
sem_list %<>% map(., ~.x + theme(plot.margin = unit(x = rep(0, 4), units = "pt")))
names(sem_list) <- type
sem_plot <- plot_grid(
  plotlist = sem_list[c("meta_bm", "bm", "meta_stab", "stab")],
  labels = "AUTO" 
  )
save_plot(
  filename = mypath("manuscript", "bef_stability", "figs",
    "sem_colin.pdf"),
  plot = sem_plot, ncol = 2, nrow = 2,
  base_width = 2.71, base_asp = 1.2 
)
#sem_plot
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "sem_mano.pdf"))
```