### 2021-02-17 manuscript/bef_stability/sem_relationship.Rmd
Produce sem indiv plot

---
title: test
---

```{r}
library(tidyverse)
library(magrittr)
library(cowplot)
library(nlme)
library(ggeffects)

mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "synchrony.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "local_network_build.R"))
theme_set(theme_cowplot())
```

# Prepare data

```{r}
myload(sem_data, community_metrics, op_analysis, dir = mypath("data"))
community_metrics %<>%
    left_join(select(op_analysis, station, opcod, year))

stab_sem_rich <- compute_stab_sem_rich(.data = as.data.frame(sem_data),
  random_effect = as.formula("~1|basin"), get_sem = TRUE)

bm_sem_rich <- compute_prod_sem_rich(.data = as.data.frame(sem_data),
  random_effect = as.formula("~1|basin"), get_sem = TRUE)
```

# Make function

```{r}
get_y_x_comb_from_model <- function(mod = NULL) {

  x_terms <- labels(terms(mod))
  all_terms <- dimnames(attr(terms(mod), "factors"))[[1]]

  y_term <- all_terms[! all_terms %in% x_terms]

  tibble(
    y = y_term, 
    x = x_terms 
  )
}

get_predict_from_model_x <- function (mod = NULL, term = NULL) {
  ggpredict(mod, terms = c(term)) %>%
    as.tibble()
}

get_predict_plot_from_model_x <- function (mod = NULL, x = NULL) {

  stopifnot(is.character(x))

  pred <- get_predict_from_model_x(mod = mod, term = x)

  y <- unique(get_y_x_comb_from_model(mod = mod)$y)

  dataset <- mod$data

  # In case of classic lm model:
  if (is.null(dataset)) {
    dataset <- mod$model
  }

  p <- dataset %>%
    ggplot(aes_string(y = y, x = x)) +
    geom_point()

  p <- p +
    geom_line(data = pred, aes(y = predicted, x = x))

  # Axis label
  label_tmp <- get_sem_var_name_replacement()
  label <- str_replace_all(label_tmp, "\n", "") 
  names(label) <- names(label_tmp)
  p +
    labs(
      x = label[x],
      y = label[y]
    )
}
```

## Test function

```{r}
test <- get_y_x_comb_from_model(stab_sem_rich[[1]])

map(test$x, ~get_predict_from_model_x(stab_sem_rich[[1]], .x))

get_predict_plot_from_model_x(stab_sem_rich[[1]], x = "log_RC1")
```

# Produce plot

```{r}
class(stab_sem_rich) <- "list"
# drop data element from the list
stab_sem_rich[[length(stab_sem_rich)]] <- NULL

ti <- tibble(model = stab_sem_rich,
  comb_term = map(model, ~try(get_y_x_comb_from_model(.x)))) %>%
  unnest(comb_term)
stab_plot <- 
  ti %>%
  mutate(
    gg = map2(x, model, ~try(get_predict_plot_from_model_x(mod = .y, x = .x)))
  )
plot_grid(plotlist = stab_plot$gg)

# Write plots:
map(stab_plot$gg,
  ~save_plot(file = mypath(
      "manuscript", "bef_stability", "result", 
      paste0(
	"stab_",
	names(.x$labels$y),
	"_",
	names(.x$labels$x),
	".png"
      )
      ), .x) 
)

mysave(stab_plot, mypath("manuscript", "bef_stability", "result"), overwrite = TRUE)
```

```{r}
class(bm_sem_rich) <- "list"
# drop data element from the list
bm_sem_rich[[length(bm_sem_rich)]] <- NULL

bm_mod <- tibble(model = bm_sem_rich,
  comb_term = map(model, ~try(get_y_x_comb_from_model(.x)))) %>%
  unnest(comb_term)
bm_plot <- 
  bm_mod %>%
  mutate(
    gg = map2(x, model, ~try(get_predict_plot_from_model_x(mod = .y, x = .x)))
  )
plot_grid(plotlist = bm_plot$gg)

# Write plots:
map(bm_plot$gg,
  ~save_plot(file = mypath(
      "manuscript", "bef_stability", "result", 
      paste0(
	"bm_",
	names(.x$labels$y),
	"_",
	names(.x$labels$x),
	".png"
      )
      ), .x) 
)
mysave(bm_plot, mypath("manuscript", "bef_stability", "result"), overwrite = TRUE)
```