### 2021-02-18 manuscript/bef_stability/fig2.Rmd
Sensivity analysis for referee two and three

---
title: "Figure 2"
---

```{r}
library(tidyverse)
library(magrittr)
library(cowplot)

mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

source(mypath("R", "misc.R"))
source_dir(mypath("R"))
```

```{r}
load(mypath("manuscript", "bef_stability", "result", "workspace.rda"))
ti <- stab_sem_rich$coefficients[,-9] %>%
  filter(P.Value <= 0.05) %>%
  mutate(
    std_est = round(Std.Estimate, 2),
    abs_std_est = abs(std_est)
  )
ti %>%
  mutate(width_arrow = abs_std_est * 0.3 / min(abs_std_est)) %>%
  select(Response, Predictor, Estimate, Std.Estimate:width_arrow)
```
```{r}
ti <- bm_sem_rich$coefficients[,-9] %>%
  filter(P.Value <= 0.05) %>%
  mutate(
    std_est = round(Std.Estimate, 2),
    abs_std_est = abs(std_est)
  )
ti %>%
  mutate(width_arrow = abs_std_est * 0.3 / min(abs_std_est)) %>%
  select(Response, Predictor, Estimate, Std.Estimate:width_arrow)
```