### 2021-02-18 manuscript/bef_stability/review_analysis.Rmd

---
title: "Analyse from review"
output: bookdown::html_document2
---

```{r, message = FALSE, results = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  comment = "#>",
  cache = TRUE,
  echo = FALSE,
  warning = FALSE,
  warnings = FALSE
)
#knitr::opts_chunk$get()
```

```{r}
mypath <- rprojroot::find_package_root_file

library(tidyverse)
library(cowplot)
library(magrittr)

source(mypath("R", "misc.R"))
source_dir(mypath("R"))

theme_set(theme_alain())

myload(sem_data, synchrony, dir = mypath("data"))
```

# Response to the comment R1-5

We check that $CV_{com, i} = \overline{CV_{sp, i}} \times \sqrt{\phi_i}$

```{r}
synchrony %>%
  ggplot(aes(y = cv_sp * sqrt(synchrony), x = cv_classic)) +
  geom_point() +
  labs(y = "CVsp * sqrt(synchrony)", x = "CV of biomass")
```

# Response to the comment R3-1

```{r}

mod <- lm(log_stab_std ~ 0 + log_cv_sp + log_sync, data = sem_data)

#summary(mod)
coefs <- broom::tidy(mod) %>%
  select(term,estimate)

coefs %<>%
  mutate(
    sd = map_dbl(term, ~sd(sem_data[[.x]])),
    coef_std = (estimate * sd) / sd(sem_data$log_stab_std)
  )

# SD de la variable de Olivier et al. (2020):  
sd(log(1 / sem_data$sync))
sd(log(1 / sem_data$cv_sp))
```

# Response to the comment R3-L231

```{r}
myload(op_analysis, dir = mypath("data"))

# station used in sem_data
op <- op_analysis %>%
  filter(station %in% sem_data$station)

## Combien de station par année
op %>%
  group_by(year) %>%
  summarise(n = n()) %>% 
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity")
```

```{r, fig.cap=""}
match_op_temp_data <- op %>%
  group_by(station) %>%
  summarise(
    year_temp = length(which(year %in% seq(2006, 2018))),
    year_temp_percent = length(which(year %in% seq(2006, 2018))) / n()
  )

match_op_temp_data %>%
  ggplot(aes(x = year_temp)) +
  geom_histogram() +
  labs(y = "Frequency", x = "Nb year of overlap between sampling and temperature time series")

match_op_temp_data %>%
  ggplot(aes(x = year_temp_percent)) +
  geom_histogram() +
  labs(y = "Frequency", x = "Percent of sampling years which overlaps with temperature time series")
```

```{r}
library(nlme)

myload(biomass_ts_sax, habitat_pressure, dir = mypath("data"))
myload(network_metrics, dir = mypath("data", "classes"))

# Get op only from 2006: 
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_bbb_start_2006 <- op_analysis_bbb %>%
  filter(year >= 2006) 
# Select station followed 10 years or more
op_analysis_bbb_start_2006 %<>%
  group_by(station) %>%
  filter(n() >= 10) %>%
  ungroup()

st_basin <- get_basin_station()

sem_data_2006 <- compute_sem_dataset_from_op(
  .op = op_analysis_bbb_start_2006,
  habitat_pressure = habitat_pressure,
  st_basin = st_basin
)

stab_sem_rich_2006 <- compute_stab_sem_rich(
  .data = sem_data_2006,
  random_effect = as.formula("~1|basin")
)
stab_sem_rich <- compute_stab_sem_rich(
  .data = sem_data,
  random_effect = as.formula("~1|basin")
)
data_sensivity_2006 <- list(
  full = sem_data,
  from_2006 = sem_data_2006 
)

nb_station <- map_int(data_sensivity_2006, ~length(unique(.x$station)))
```

```{r stab}
stab_sensivity_2006 <- 
  map(data_sensivity_2006,
    ~compute_stab_sem_rich(
      .data = .x,
      random_effect = as.formula("~1|basin")
    )
  )

sem_stab_results_2006 <- map(stab_sensivity_2006,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_stab_results_2006) <- str_to_lower(colnames(sem_stab_results_2006))

p_sens_stab <- plot_sensivity_analysis(
  .data = sem_stab_results_2006,
  inverse_label_order = TRUE, 
  y_lim = c(-1.5, .75)
  ) +
scale_color_manual(
  values = c("full" = "red", "from_2006" = "blue"),
  labels = c("Full dataset", "Dataset from 2006"),
  breaks = c("full", "from_2006") 
  ) +
  guides(color = guide_legend(direction = "horizontal", nrow = 1)) +
  theme(legend.position = "bottom", plot.margin = unit(c(0, 0, 0, 25), units = "mm"))
```

```{r}
bm_sensivity_2006 <- 
  map(data_sensivity_2006,
    ~compute_prod_sem_rich(
      .data = .x,
      random_effect = as.formula("~1|basin")
    )
  )
sem_bm_results_2006 <- map(bm_sensivity_2006,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_bm_results_2006) <- str_to_lower(colnames(sem_bm_results_2006))

p_sens_bm <- plot_sensivity_analysis(
  .data = sem_bm_results_2006,
  inverse_label_order = TRUE,
  y_lim = c(-1.5, .75)
  ) +
scale_color_manual(
  values = c("full" = "red", "from_2006" = "blue")
  ) +
  guides(color = guide_legend(direction = "horizontal", nrow = 1)) +
  theme(legend.position = "bottom", plot.margin = unit(c(0, 0, 0, 25), units = "mm"))
```

```{r start-sens, fig.dim = c(14, 10), fig.cap="Sensivity of standardized estimate to the station starting date."}
legend_b <- get_legend(p_sens_stab)

p <- plot_grid(
  p_sens_bm %<>% + theme(legend.position = "none", plot.margin = unit(c(0,0,0,27), units = "mm")),
  p_sens_stab %<>% + theme(legend.position = "none", plot.margin = unit(c(0,0,0,27), units = "mm")),
  legend_b, nrow = 3,
  rel_heights = c(1, 1, .1),
  labels = c("Total biomass", "Biomass stability")
)
p
```

## L180-182

```{r}
library(nlme)

compute_bm_sem_rich_var <- function(.data, rich_var = "log_rich_tot_std", random_effect = "~ 1 | basin", get_sem = FALSE) {

  formula_list <- list(
    rich = paste0(rich_var, " ~ log_RC1 + log_RC2"),
    ct = paste0("ct ~ log_RC1 + log_RC2 + ", rich_var),
    t_lvl = paste0("t_lvl ~ log_RC1 + log_RC2 + ", rich_var),
    bm = paste0("log_bm_std ~ ", rich_var," + ct + t_lvl + log_RC1 + log_RC2")
  )
  model_list <- map(formula_list, function(x)  {
    temp_model <- nlme::lme(as.formula(x), random = ~ 1 | basin, data = .data)
    temp_model$call$fixed <- as.formula(x)
    return(temp_model)
  }
    )

  corsem <- piecewiseSEM::as.psem(model_list)

  if (get_sem) {
    return(corsem)
  }
  output <- summary(corsem, .progressBar = F)

  return(output)
}
compute_stab_sem_rich_var <- function(.data, rich_var = "log_rich_tot_std", random_effect = "~ 1 | basin", get_sem = FALSE) {

  formula_list <- list(
    rich = paste0(rich_var, " ~ log_RC1 + log_RC2"),
    ct = paste0("ct ~ log_RC1 + log_RC2 + ", rich_var),
    t_lvl = paste0("t_lvl ~ log_RC1 + log_RC2 + ", rich_var),
    sync = paste0("log_sync ~ ", rich_var, " + ct + t_lvl + log_RC1 + log_RC2"),
    cvsp = paste0("log_cv_sp ~ ", rich_var," + ct + t_lvl + log_RC1 + log_RC2") 
    )

  model_list <- map(formula_list, function(x)  {
    temp_model <- nlme::lme(as.formula(x), random = ~ 1 | basin, data = .data)
    temp_model$call$fixed <- as.formula(x)
    return(temp_model)
  }
    )

  ## Add linear model
  model_list <- c(
    model_list,
    stab = list(lm(log_stab_std ~ log_cv_sp + log_sync, data = .data))
  )

  corsem <- piecewiseSEM::as.psem(model_list)

  if (get_sem) {
    return(corsem)
  }
  output <- summary(corsem, .progressBar = F)

  return(output)
}
sem_data %<>%
  mutate(log_rich_std = log(rich_std_med))
```

```{r}
rich <- names(sem_data)[str_detect(names(sem_data), "log_rich")]
```

```{r}
bm_sens_rich <- map(rich, 
  ~compute_bm_sem_rich_var(
    .data = sem_data,
    rich_var = .x, random_effect = "~ 1 | basin",
    get_sem = FALSE)
)
names(bm_sens_rich) <- rich

sem_bm_results_rich <- map(bm_sens_rich,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value) %>%
  mutate_at(c("Response", "Predictor"), ~ifelse(str_detect(., "rich"), "richness", .))

colnames(sem_bm_results_rich) <- str_to_lower(colnames(sem_bm_results_rich))

cols_rich <- c(
  "log_rich" = "red",
  "log_rich_std" = "blue",
  "log_rich_tot_std" = "darkgreen",
  "log_rich_tot" = "orange"
)
label_rich <- c(
  "log_rich" = "Log median richness",
  "log_rich_std" = "Log median of standardized richness",
  "log_rich_tot_std" = "Log standardized total richness",
  "log_rich_tot" = "Log total richness"
)

p_sens_bm_rich <- plot_sensivity_analysis(
  .data = sem_bm_results_rich,
  inverse_label_order = TRUE, scale_color = FALSE,
  y_lim = c(-1.5, .75)
  ) +
scale_color_manual(
  values =  cols_rich,
  labels = label_rich,
  breaks = names(label_rich)) +
  guides(color = guide_legend(direction = "horizontal", nrow = 1)) +
  theme(legend.position = "bottom", plot.margin = unit(c(0, 0, 0, 25), units = "mm"))
```

```{r}
stab_sens_rich <- map(rich, 
  ~compute_stab_sem_rich_var(
    .data = sem_data,
    rich_var = .x, random_effect = "~ 1 | basin",
    get_sem = FALSE)
)
names(stab_sens_rich) <- rich

sem_stab_results_rich <- map(stab_sens_rich,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value) %>%
  mutate_at(c("Response", "Predictor"), ~ifelse(str_detect(., "rich"), "richness", .))

colnames(sem_stab_results_rich) <- str_to_lower(colnames(sem_stab_results_rich))

cols_rich <- c(
  "log_rich" = "red",
  "log_rich_std" = "blue",
  "log_rich_tot_std" = "darkgreen",
  "log_rich_tot" = "orange"
)
label_rich <- c(
  "log_rich" = "Log median richness",
  "log_rich_std" = "Log median of species richness density",
  "log_rich_tot_std" = "Log species richness density",
  "log_rich_tot" = "Log total richness"
)

p_sens_stab_rich <- plot_sensivity_analysis(
  .data = sem_stab_results_rich,
  inverse_label_order = TRUE,
  scale_color = FALSE, y_lim = c(-1.4, .75)) +
scale_color_manual(
  values =  cols_rich,
  labels = label_rich,
  breaks = names(label_rich)) +
  guides(color = guide_legend(direction = "horizontal", nrow = 1)) +
  theme(legend.position = "bottom", plot.margin = unit(c(0, 0, 0, 25), units = "mm"))
```

```{r rich-sens, fig.dim = c(14, 10), fig.cap="Sensivity of the standardized estimates to species richeness specification for the two structural equation model."}
legend_b <- get_legend(p_sens_stab_rich)
p <- plot_grid(
  p_sens_bm_rich %<>% + theme(legend.position = "none", plot.margin = unit(c(0,0,0,27), units = "mm")),
  p_sens_stab_rich %<>% + theme(legend.position = "none", plot.margin = unit(c(0,0,0,27), units = "mm")),
  legend_b, nrow = 3,
  rel_heights = c(1, 1, .1),
  labels = c("Total biomass", "Biomass stability")
)
p
```

The choice of richness variable would

# Response to comment R3-Section 5.3

```{r}

```