### 2021-02-19 manuscript/bef_stability/table/hypothesis_sem.tex
Minor edits

\begin{tabular}{| m{7cm}| m{2cm}| m{4cm}|}
    \hline
    Link & hypothesis & reference\\
    \hline
    PCA1 Avg stream size $\to$ Species richness & (+) &
    \citet{post_ecosystem_2000, doi_resource_2009, mchugh_dual_2010} \\
    \hline
    PCA2 Avg temperature $\to$ Species richness & (+/-) & \citet{ogorman_unexpected_2017,gauzens_biodiversity_2020} \\
    \hline
    PCA1 Avg stream size $\to$ Connectance & ? & NA\\
    \hline
    PCA2 Avg temperature $\to$ Connectance & ? & NA\\
    \hline
    Species richness $\to$ Connectance & (-) & \citet{dunne_network_2006} \\
    \hline
    PCA1 Avg stream size $\to$ Avg trophic level & (+) & \citet{mchugh_dual_2010}\\
    \hline
    PCA2 Avg temperature $\to$ Avg trophic level & (+/-) & \citet{ogorman_unexpected_2017}\\
    \hline
    Species richness $\to$ Avg trophic level & ? & NA\\
    \hline
    Species richness $\to$ Synchrony & (-) & \citet{thebault_trophic_2005, loreau_species_2008, thibaut_understanding_2013} \\
    \hline
    Connectance $\to$ Synchrony & (+) & \citet{teng_dynamics_2004}\\
    \hline
    Avg trophic level $\to$ Synchrony & (+) & \citet{raimondo_interspecific_2004}\\
    \hline
    PCA1 Avg stream size $\to$ Synchrony & (-) & NA\\
    \hline
    PCA2 Avg temperature $\to$ Synchrony & (+) & \citet{loreau_species_2008}\\
    \hline
    Species richness $\to$ CVsp & (+) & \citep{tilman_biodiversity_1996} \\
    \hline
    Connectance $\to$ CVsp & NA & NA\\
    \hline
    Avg trophic level $\to$ CVsp & (-) & \citet{shanafelt_stability_2018}\\
    \hline
    PCA1 Avg stream size $\to$ CVsp & (-) & NA\\
    \hline
    PCA2 Avg temperature $\to$ CVsp & (+) & \citep{ogorman_unexpected_2017;tabi_warming_2019}\\
    \hline
    CVsp $\to$ Biomass stability & (-) & \citet{thibaut_understanding_2013}\\
    \hline
    Synchrony $\to$ Biomass stability & (-) & \citet{thibaut_understanding_2013}\\
    \hline
    Species richness $\to$ Total biomass & (+) & \citet{schneider_animal_2016}\\
    \hline
    Connectance $\to$ Total biomass & NA & \citet{wang_biodiversity_2018}\\
    \hline
    Avg trophic level $\to$ Total biomass & (+) & \citet{wang_biodiversity_2018}\\
    \hline
    PCA1 Avg stream size $\to$ Total biomass & (+) & \citet{post_ecosystem_2000, doi_resource_2009, @allan_stream_2007}\\
    \hline
    PCA2 Avg temperature $\to$ Total biomass & (+/-) & \citep{ogorman_unexpected_2017}\\
    \hline
\end{tabular}