### 2021-02-19 manuscript/bef_stability/table/hypothesis_table.Rmd

---
title: "Make a table for hypothesis"
output: 
  bookdown::pdf_document2:
    citation_package: natbib
biliography: references.bib
---

```{r, message = FALSE, results = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  comment = "#>",
  cache = TRUE,
  echo = FALSE,
  warning = FALSE,
  warnings = FALSE
)
#knitr::opts_chunk$get()
```

```{r}
mypath <- rprojroot::find_package_root_file

library(tidyverse)
library(cowplot)
library(magrittr)
library(kableExtra)

source(mypath("R", "misc.R"))
source_dir(mypath("R"))

```

```{r}
sem_hyp <- tibble(
  link = "Species richness \\to connectance",
  pred = "Connectance is negatively linked to species richness",
  ref = "\\citep{dunne_network_2006}"
)
sem_hyp <- tibble(
  link = "Species richness \\to connectance",
  pred = "Connectance is negatively linked to species richness",
  ref = "@dunne_network_2006"
)

sem_hyp %>%
  kable(format = "latex")

```
(ref:studies-table-ref1) Bogaert *et al.* (2011) [@dunne_network_2006]

```{r}
sem_hyp <- tibble(
  link = "Species richness to connectance",
  pred = "Connectance is negatively linked to species richness",
  ref = ""
)
library(nlme)
myload(sem_data, dir = mypath("data"))
stab_sem_rich <- compute_stab_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))
bm_sem_rich <- compute_prod_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))
resp_pred <- map_dfr(list(stab_sem_rich, bm_sem_rich), ~.x$coefficients) %>% 
  distinct(Predictor, Response) %>%
  select(pred = Predictor, resp = Response)

var_replacement <- str_replace_all(get_sem_var_name_replacement(),
  c("\n" = "", "\\s{2}" = " "))
names(var_replacement) <- names(get_sem_var_name_replacement())

ta <- resp_pred %>%
  mutate_all(~str_replace_all(., var_replacement)) %>%
  unite(Link, pred:resp, sep = " xto ") %>%
  mutate(hypothesis = NA, reference = NA)

ti <- ta %>%
  kable(format = "latex")

test <- str_replace_all(ti, "xto", "$\to$")
class(test) <- class(ti) 
attr(test, "format") <- attr(ti, "format")

write(ti, file = mypath("manuscript", "bef_stability", "hypothesis_sem_empty.tex"))
```