### 2021-04-26 manuscript/bef_stability/appendixB.Rmd
Updates

---
title: Appendix B of Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
nocite: |
    @post_ecosystem_2000| @doi_resource_2009| @mchugh_dual_2010|
    @ogorman_unexpected_2017| @gauzens_biodiversity_2020| @dunne_network_2006|
    @mchugh_dual_2010| @doi_resource_2009| @allan_stream_2009|
    @petchey_predicting_2010| @winemiller_must_1989
csl: ecology-letters.csl
---

```{r, echo = FALSE, message = FALSE, collapse = FALSE}

mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

load(file = mypath("manuscript", "bef_stability", "result", "workspace.rda"))

knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
library(nlme)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))
source(mypath("R", "cleaning_methods.R"))

theme_set(theme_alain())
```

\beginsupplement

```{r, echo = FALSE}
package <- c("openSTARS", "SSN", "nlme", "piecewiseSEM")
#purrr::map(package, ~ toBibtex(citation(.x)))

#toBibtex(citation())

#print(cm, bibtex = TRUE)
# Load manuscript results
```

# Data collection and preparation

## Fishing

### Protocol

The monitoring of the fish communities occurs in a variety of stream size, shape
and depth. The stream characteristics constrain the type of protocol that can be
applied. Small streams (depth < 0.7m across the entire width) are sampled with
the complete protocol. The protocol consists of placing anodes across the entire
width of the stream. The fishermen made from one to three rows of electric
fishing. We selected only the data from the first row to homogenize the sampling
effort. In deeper and wider streams (depth >= 0.7m), the sampling was partial
over the bank and the channel and performed by boat and by foot (along the
bank). We selected only the sampling done over the bank because the channel was
not systematically sampled. Two types of partial protocols takes place: partial
over bank and partial by points, which are respectively continuous and discrete.
Furthermore, in the "partial by point" protocol, the channel was sometimes
sampled as well. We selected only the "partial by point" sampling in which all
the sampling points were done over the bank to homogenize the sampling. Then, we
merged partial over bank and partial by point protocols. Some sampled stations
were located in stream with intermediate characteristics where both protocols
were alternatively used depending on the state of the stream at the time of
sampling. For these sites, we only kept the samples that had been carried out
with the most used protocol.

Heterogeneity in the sampling date may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in the sampling date, only one sampling per year was retained if
there was more in a given year, the one with that had the most species and then
the most individuals. Finally, most of the fish sampling occurred in September
(Median: mid-August) and the median standard deviation of the sampling month was
inferior to one (0.7) within stations, i.e. 3 weeks. To further reduce sampling
heterogeneity, we only kept the sampling in which the length of stream sampled
along the stream was inside the median length sampled more or less 30%.

### Batches definition and body size measurement

Once the fishes have been caught, they are divided into different batches (Fig.
\@ref(fig:p-body-size)). Each batch contains only one species but there may be
several batches for one species. The body size of the fishes is measured
differently for each type of batch.  There are four types of batches: N, I, S/L
and G. The batches N and I contain the biggest individuals. In the N and I
batches, the body size of all the individuals is recorded. The batches S/L
contain fishes of intermediate sizes.  Inside each S/L batch, the operator tries
to make batches as homogeneous as possible in terms of body size. Then, the body
size of a subsample of at least 20 individuals is recorded. The G batches
contain individuals of small body size, with care that body size is as
homogeneous as possible in a given batches.  Then body size of the shortest and
the longest individuals are recorded.

## Body size inference

```{r}
myload(fish_length, lot_id_opcod, dir = mypath("data-raw", "fishing_op_build"))

fish_length %<>%
  left_join(lot_id_opcod) %>%
  rename(length = fish) %>%
  select(opcod, species, length)

length_sem_analysis <- fish_length %>% 
  filter(opcod %in% op_analysis_bbb$opcod)
rm(fish_length, lot_id_opcod)
nrow_fish_length <- nrow(length_sem_analysis)
na_length <- sum(is.na(length_sem_analysis$length))
```

The instructions for body size inference were guided by the recommendations of
the OFB agency, which has defined the sampling protocols, supervised the data
collection and performed analysis for multiple purpose including stream
management.

```{r, cache = FALSE}
body_size_caption <- paste0(
  "Body size inference of the fishes. ",
  "(1) Sampled fishes were attributed to batch type according to their size and
  the number of individuals. A batch contains one species only. (2) In the batch
  type N and I, the body size of all the individuals was recorded. In the
  batches, S/L, the body size of a subsample of the individuals was recorded.
  Finally, only the body size of the smallest and of the largest individual were
  recorded in the G batches. ",
  "Then, (3) we inferred the body size distribution in the S/L and G batches by
  assuming a normal distribution. The mean and the standard
  deviation were estimated with the measured subsample."
) 
```

```{r p-body-size, cache = FALSE, fig.cap=body_size_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "length_inference.pdf"))
```

We inferred the body size of all the unmeasured individuals using the batch
records. There was no need for inference for the N and I batches since the body
size of all individuals was recorded (Fig. \@ref(fig:p-body-size)). The body
size ($l$) of the fishes in the S/L and G was inferred under the assumption that
body size distribution follows a normal law ($l ~ N(\mu, \sigma)$).  For an S/L
batch, the mean $\mu$ and the standard deviation $\sigma$ were simply the
average and the standard deviation of the body size of the subsample recorded.
For a G batch, the average was approximated by $\hat{\mu} = (l_{min} + l_{max})
2^{-1}$ and the standard deviation by $\hat{\sigma} = (l_{min} + l_{max})
4^{-1}$. With the estimations of the mean and the standard deviation, the number
of individuals in a batch and the assumption of normal distribution, we
simulated the body size distribution in S/L and G batches. To avoid simulation
of unrealistic body size, we used a truncated normal distribution bounded at
[0.05; .95] for the density of probability of the S/L batches, and bounded at
[$l_{min}$; $l_{max}$] for the G batches. Body size distribution were generated
using the R function `truncdist::rtrunc`.

Batches that did not meet certain conditions were excluded. First, we also
excluded batches for which the total number of individuals was not recorded. We
also excluded G batches for which $l_{min} \geq l_{max}$ or if
at least one of the measurement was missing. Additionally, a given G batch was
excluded when the number of individuals fell below five, as G batches normally
contained more individuals. We finally excluded S/L batches for which the
number of measured individuals was inferior to 10. Finally, in the dataset used
in the analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

## Community data

```{r}
source(mypath("R", "cleaning_methods.R"))
removed_ind <- filter(length_sem_analysis, species %in% sp_to_remove())
removed_sp <- unique(removed_ind$species)
```

The rare fish species, species from lake, crayfish species, and migratory
species were excluded. In total, we removed `r nrow(removed_ind)` over 
`r nrow(length_sem_analysis)` individuals and `r length(removed_sp)` species.

Some species were identified based on their morphotype, but they
belong to the same species based on fecundation criteria. We merged those
morphotype species into one species. There was
`r length(unique(names(sp_to_replace())))` species to merge which belonged to 
`r length(unique(sp_to_replace()))` species.

### Metaweb

#### Ontogenic diet

Diet information were obtained from fishbase.org and literature (Appendix A, Table S2). The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had a strict piscivorous
life stage. Fish species were then largely omnivorous.

```{r food, eval = FALSE}
onto_shift_cap <- paste0(
  "Ontogenic diet shift for each fish species.", 
  "Species had between two and three life stages, bounded by body size (in mm).
  A life stage of a given determine its diet. The fourth column to the last one
  correspond to this respective resource: light, detritus, biofilm,
  phytobenthos, macrophages, phytoplankton, zooplankton,  zoobenthos and
  fishes."
)
table_onto_diet <- fish_diet_shift %>%
  rename(`Life stage` = stage) %>%
  mutate(
    min = paste0("[", size_min),
    max = paste0(size_max, "]")
  ) %>%
  unite("Size interval", min:max, sep = ";") %>%
  select(species, `Life stage`, `Size interval`, light:ref_ods)

kable(table_onto_diet,
  format = "latex",
  booktabs = T,
  label = paste0("onto-diet"),
  caption = onto_shift_cap
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

```{=latex}
%\includepdf[pages={2-7}]{figs/supplementary_ref_willem.pdf}
```

#### Define trophic interactions

```{r, cache = FALSE}

trophic_building_caption <- paste0( "Building of the metaweb. (1) Each
  species are divided in nine body size classes. We gathered information about
  the food diet of each fish species and how food diet varies according to
  ontogeny. (2) The life stage of a trophic species was determined with the
  median of the body size class. A trophic link was set between the trophic
  species and the resources associated to this life stage. (3) For a piscivorous
  trophic species, the trophic links between fishes were defined according to
  the predators/prey body size ratio. The predation window of the trophic
  species ranged from 3% to 45% of the median of the body size range. A trophic
  link between the predator trophic species and every trophic species whose
  midpoint of bodysize range was included in the predation window.")

```

```{r p-net-inf, cache = FALSE, fig.cap=trophic_building_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "network_inference.pdf"))
```

```{r}
myload(metaweb_analysis, dir = mypath("data"))
species_in_metaweb <- metaweb_analysis$species

length_sem_analysis %<>%
  filter(!species %in% sp_to_remove()) %>%
  mutate(
    species = str_replace_all(species, sp_to_replace())
  )

removed_ind_metaweb <- length_sem_analysis %>%
  filter(!species %in% species_in_metaweb)
removed_sp_metaweb <- unique(removed_ind_metaweb$species)
```

```{r, cache = FALSE}
metaweb_caption <- paste0(
  "The metaweb describes the trophic interactions between all the 412 trophic species, each is
  represented by a node and each color represents a species (S = 45). The seven
  resource nodes (7) are drawn in grey. The species names corresponding to the
  three digit codes are found in Table S1."
)
```

```{r metaweb, cache = FALSE, fig.cap=metaweb_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "metaweb.png"))
```

The metaweb describes the potential trophic interactions between the
($`r ncol(metaweb_analysis$metaweb) - length(metaweb_analysis$resource)`$) trophic species.
The metaweb also included seven resource nodes which were present in the diet of the trophic
species namely detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos (Fig. \@ref(fig:metaweb)).

Trophic interactions between the nodes of the metaweb were determined according to
the diet and the body size of the trophic species. `r length(removed_sp_metaweb)` 
species were removed because information about their diet was missing. It
represented  `r nrow(removed_ind_metaweb)` individuals over
`r nrow(length_sem_analysis)` individuals.

The life stage to which a trophic species belongs was determined by the median of
the body size range of the trophic species, each life stage being delimited by a body
size range. In turn, this trophic species has trophic interactions with the
resource nodes of its life stage. The trophic interactions between a
predator and a prey were determined by predator-prey body size ratio
[@brose_predator_2019; @poisot_structure_2016; @schneider_body_2012]. For a
given trophic species, we first determined if its body size range overlap
with a piscivorous life stage (Fig. \@ref(fig:p-net-inf)). We were then more
liberal for piscivory determination than for fish-resource interactions, to
avoid missing potential interactions which were the main interactions in our
community. If the trophic species was piscivorous, we determined its prey size
range from 3% to 45% of the median of its size range. A trophic interaction was
set for each trophic species whose median of its size class fell into the prey
range of the predator trophic species. To sum up, the metaweb contains feeding
interactions (1) between trophic species, (2) between trophic species and
resource nodes and (3) between resource nodes.

#### Local trophic network

The local trophic networks (i.e. one network by
sampling) were obtained by determining the trophic species of each fish
individual in the community. According to the trophic species present in a given
community, we set the trophic interactions between them according to the metaweb. 
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species.

## Environmental variables

The procedure to obtain altitude, slope, width and depth of the streams are fully
explained in main text. Because those measurements were collected during the
fishing protocol, they were easier to manage in the analysis. Then, we detailed
in the following sections how water temperature, water flow and Biological
Oxygen Demand (BOD) variables were obtained. The goal was to obtain an annual
mean for each variable, to match the yearly resolution of the fish monitoring.
Furthermore, those variables are measured at different locations than the fish
monitoring.

### Preparation

Water temperature, water flow and BOD variables were obtained from the [naiades
database](http://www.naiades.eaufrance.fr/) and were respectively collected at
an
hourly, daily and monthly frequency. Water temperature and flow are recorded by
automated sensors while BOD is collected manually because its measurement
requires laboratory analysis. BOD data collection is coordinated by the Agence
de l'eau which realized the analysis itself or order it to a company.

```{r, cache = FALSE}
pressure_cap <- paste0(
  "Map of the environmental variables obtained for the 99 sites used in the
  analysis. Slope is expressed in degree, altitude in meter, distance from
  source in kilometer, Strahler order has no units, average stream width and
  depth are expressed in meter, Biological Oxygen Demand (BOD) is expressed in
  mg of dissolved oxygen, water flow is expressed in cube meter per second and
  water temperature is expressed in Celsius degree.")
```

```{r, cache = FALSE, fig.height = 20/2.54, fig.cap = pressure_cap}
pca_pressure <- habitat_press
myload(fourier, dir = mypath("data-raw", "polluants"))
myload(flow_temp_metrics, station_analysis, the_8_hydrologic_basin, dir = mypath("data"))

sf_pressure <- station_analysis %>% 
  filter(id %in% op_analysis_bbb$station) %>%
  select(id) %>%
  left_join(rename(pca_pressure, id = station)) %>%
  na.omit()

p <- ggplot() +
  geom_sf(data = the_8_hydrologic_basin)

var_to_plot <- colnames(pca_rotated$normal$tab)

var_replacement <- c(
  "slope" = "Slope",
  "alt" = "Altitude",
  "d_source" = "Distance from source",
  "strahler" = "Strahler order",
  "width_river_mean" = "Average stream width",
  "avg_depth_station_mean" = "Average stream depth",
  "width_river_cv" = "CV of stream width",
  "avg_depth_station_cv" = "CV of stream depth",
  "DBO_med" = "Average BOD",
  "flow_med" = "Average flow",
  "temperature_med" = "Average temperature",
  "DBO_cv" = "CV of BOD",
  "flow_cv" = "CV of flow",
  "temperature_cv" = "CV of temperature"
)

map_pressure <- map(var_to_plot,
~p +
  geom_sf(data = sf_pressure, aes_string(color = .x), size = .5) +
  viridis::scale_color_viridis() +
  labs(title = str_replace_all(.x, var_replacement)) +
  theme(
    legend.title = element_blank(),
    text = element_text(size = 8),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
 guides(color = guide_colourbar(barwidth = 0.5, barheight = 4))
)

plot_grid(plotlist = map_pressure, ncol = 3)
```

To get rid of seasonal effects of temperature, we applied a moving window
averaging to each time series. The moving window was applied on daily average
temperature. Then we took the annual mean of each year and each variable
(i.e. one value by year and by station).

Several characteristics of the streams make the interpolation more complicated
than along terrestrial surface. Interpolations in terrestrial areas are roughly
based on euclidean distance while euclidean distance is complicated to consider
in streams because it had to follow the path of the streams. Furthermore, two
streams can be close by but not connected and thus the value of their variables might 
not be related or only a few. For those reason, we used stream spatial network
methods to interpolate water temperature, water flow and BOD.

We reconstructed the stream network and the connection between stations with the
`openSTARS` R package [@kattwinkel_preparing_2020]. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(IGN)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network [@pella_reseau_2012].

The interpolations were carried out by hydrographic basin due to convergence
issues when the interpolation was carried out at the scale of the whole
France. We had to find a trade-off between the number of data (to keep enough
high to perform the interpolation) and the level of dataset splitting. We then
merged some hydrographic basin to end up with four regions of interpolation:
West, East, South and North.

Then we modelled each variable as dependant only on the covariation between
sites and taking in account stream network. To model the autocovariance between
sites, we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their predictive performance. The interpolations
were done with the `SSN` R package [@isaak_applications_2014; @hoef_ssn:_2014].

### Performance and post check

```{r, cache = FALSE}
cv_bias_caption <- paste0(
  "Bias of the interpolation from Leave-One-Out Cross Validation (LOOCV). "
)
```

```{r, eval = TRUE}
myload(cv_press_interp_mv_avg,
  dir = mypath("data-raw", "polluants"))
myload(cv_temp_press_interp_mv_avg,
  dir = mypath("data-raw", "naiades_temperatures"))
myload(cv_flow_press_interp_mv_avg,
  dir = mypath("data-raw", "flow"))
dbo_cv <- cv_press_interp_mv_avg %>%
  filter(parameter == "dbo5")

cv_dbo_temp_flow <- map2_dfr(
  list(dbo = dbo_cv, temp = cv_temp_press_interp_mv_avg, flow = cv_flow_press_interp_mv_avg),
  c("BOD", "Temperature", "Flow"),
  function(x, y) {
    x %>%
      select(basin, year, bias, std.bias, RMSPE, std.MSPE) %>%
      group_by(basin) %>%
      summarise_if(is.numeric, list(avg = ~mean(., na.rm = TRUE), sd = ~sd(., na.rm = TRUE))) %>%
      mutate(Variable = y) %>%
      select(Variable, basin, everything()) %>%
      ungroup
  })

```
```{r, cache = FALSE}

cv_dbo_temp_flow %>%
  mutate_if(is.double, ~signif(., 2)) %>%
  kable(
    align = "c",
    format = "latex",
    booktabs = T,
    caption = paste0("Results of the cross-validation for the interpolation of
      BOD, temperature and flow. Average over 4 interpolations are shown. The
      cross-validation were performed with the SSN R package. bias: average
      predicted error, std.biais: bias standardized, RMSPE: Root Mean-Squared
      Prediction Error, std.MSPE: standardized Mean-Squared Prediction Error.") ) %>%
    collapse_rows(columns = c(1, 2), valign = "top") %>%
    kable_styling(latex_options =c("striped", "scale_down"))
```

The performance of the interpolations was evaluated with leave one out
cross-validation (already implemented in the SSN package), i.e. the value of the
interpolated variable of each station was predicted after removing it from the
dataset.

When predicted values were below the minimum value or above the maximum value
recorded in the dataset for flow, temperature and BOD, we replaced the
predicted value by respectively the minimum and the maximum value found in the
dataset. We ran the interpolation four times, and we kept the predicted value
which had the least standard deviation.

# Data analysis

## Temporal series classification

The CV of biomass over time can cover multiple temporal dynamics, including
increases, stable, decreases and a mix of them. Theoretically, however, the CV
computed to quantify stability is often represented with the assumption that the
overall dynamic of the biomass varies around an equilibrium
[@kefi_advancing_2019]. In order to approach the condition of equilibrium, we kept only the
stations that have met this criteria. To characterise the biomass temporal
dynamics, we used the symbolic aggregated approximation method (SAX), which is
commonly used in time series analysis and implemented in the `jmotif` R package
[@senin_grammarviz_2018].  This method classifies time series rather than
quantifies trends and so is generic enough to encompass a diversity of temporal
trajectories. The classification necessitates three steps. First, the time
series are normalized (i.e. substituting the mean and divide by the standard
deviation). The second step is to reduce the dimensionality of the time series
by a Piecewise Aggregation Approximation (PAA) method. Each time series is
divided in $N$ frames of equal size that allows to reduce the dimensionality
of the time series. For each frame, we simply kept the mean value of the time
series. The last step is to qualify each frame into string. Because values of
normalized time series typically follows a normal distribution ($N(0,1)$), a
breakpoint can be computed by dividing the area under the normal distribution
in equal parts. The algorithm attributes an alphabetic letter to each part from
the lowest to the highest ($a$, $b$, $c$, ...). We set three breakpoints, so
$a$, $b$ and $c$.  Finally, the trajectory of each time series is classified as
a combination of $a$, $b$, $c$ letters. $abc$ and $cba$ time series represent
respectively monotonic increasing and decreasing trajectories. For data
analysis, we kept only the $bbb$ which represent time series that showed no
temporal tendencies, with respect to our methodology.

## PCA over environmental variables

```{r eig-plot}
rotated_eigen_values <- 
  tibble(
    name = paste0("RC", seq(1,14)),
    type = c(rep("Kept", 2), rep("No kept", 14 - 2)),
    value = pca_rotated$rotated$values
  )
eigen_rotated_plot <-
  ggplot(rotated_eigen_values,
    aes(x = reorder(name, -value) , y = value)) +
  geom_col(aes(fill = type)) +
  scale_fill_discrete(name = "Status") +
  labs(x = "Rotated component", y = "Eigen values") +
  theme(legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank())
```

```{r contrib-pca, cache = FALSE}
#https://stats.stackexchange.com/a/143949
caption_contrib_pca <- paste0(
  "Loadings of variable to rotated components of the PCA.
  Loadings are the coefficients which determine the components of the PCA as
  linear combinations of the variables.
  "
)

loadings_pca <- matrix(
  as.matrix(round(pca_rotated$rotated$Structure, 2)),
  nrow = 14,
  dimnames = dimnames(pca_rotated$rotated$Structure)
  ) %>%
  as_tibble %>%
  mutate(
    Variable = str_replace_all(
      dimnames(pca_rotated$rotated$Structure)[[1]],
      get_pca_var_name_replacement())
  ) %>%
  select(Variable, RC1, RC2)
kable(
  x = loadings_pca,
  format = "latex",
  booktabs = T,
  label = "contrib-pca",
  caption = caption_contrib_pca
  ) %>%
kable_styling(latex_options = c("striped"))
```

```{r, eval=FALSE, cache = FALSE}
corr_pca_caption <- paste0(
  "Spearman correlation between environmental variables used in the PCA.
  The variables have been scaled before to compute the correlation."
)

pca_var_correlation <- pca_pressure %>%
  select(-station) %>%
  #mutate_all(scale) %>%
  as.matrix() %>%
  cor(., method = "spearman")

pca_var_correlation %<>% round(., 2)
dimnames(pca_var_correlation) %<>%
  map(., ~str_replace_all(.x, get_pca_var_name_replacement()))

kable(
  x = pca_var_correlation,
  format = "latex",
  booktabs = T,
  label = "corr-pca",
  caption = corr_pca_caption
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r, cache = FALSE}
prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))

tot_var_expl_pca <- round(sum(pca_rotated$rotated$Vaccounted["Proportion Var", ]) * 100)

pca_caption <- paste0(
  "Eigenvalues distribution, we kept the first two axis for the data analysis." 
  )
```

```{r pca-rotated, cache = FALSE, fig.cap = pca_caption}
p1 <- my_pca_plot(xaxis = "RC1", yaxis = "RC2", ctb_thld = .4)
#p2 <- my_pca_plot(xaxis = "RC3", yaxis = "RC4", ctb_thld = .4)
#p3 <- my_pca_plot(xaxis = "RC4", yaxis = "RC5", ctb_thld = .4)

pca_plot <- plot_grid(eigen_rotated_plot, nrow = 1)
save_plot(
  filename = mypath("manuscript", "bef_stability", "figs", "pca_plot.pdf"),
  pca_plot,
  nrow = 1, ncol = 2,
  base_height = 6/2.54, base_asp = 1.6, base_width = NULL 
  )
pca_plot
```

```{r}
print_prop_var_pca <- prop_var_expl_pca[order(names(prop_var_expl_pca))]
```

```{r }
rc2 <- tibble(
  station = pca_pressure$station,
  alt = pca_pressure$alt,
  temperature_med = pca_pressure$temperature_med,
  value = pca_rotated$rotated$scores[,"RC2"]
)

rc2_sf <- station_analysis %>%
  select(station = id) %>%
  filter(station %in% rc2$station) %>%
  left_join(rc2, by = "station")

caption_rc2_map <- paste0(
  "Map of the station values for the second axis of the rotated PCA."
)
```

```{r rc2-map, eval = FALSE, fig.cap = caption_rc2_map}
p +
  geom_sf(data = rc2_sf, aes(color = log(value + abs(min(value)) + 1 ))) +
  viridis::scale_color_viridis() +
  labs(title = "RC2 (Log)") +
  theme(legend.title = element_blank())
```

```{r pca-two-axis, eval = FALSE}
two_axis <- compute_rotated_pca(.data = pca_pressure, naxis = 2)
two_axis$rotated$Vaccounted

cor(two_axis$rotated$score[,c("RC1", "RC2")],
  pca_rotated$rotated$score[,c("RC1", "RC2")])
```

```{=latex}
\pagebreak
```

# Statistical analysis

```{r, cache = FALSE}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c("Log base 10", "Log base 10 and inversion", "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  description = c("linked to stream size",
    "linked to average temperature and BOD",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level",
    "Temporal synchrony in the dynamics of population biomass", "Average
    variability of population biomass", "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "The description of the variables used in the structural equation models."
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

\begin{table}
\caption{\label{tab:hyp-sem} Presentation of the links 
of the Structural Equation Models associated with the references supporting the links.}
\begin{center}
\begin{tabular}{ m{7cm}| m{1cm}| m{6cm}}
    \hline
    Link & sign & reference\\
    \hline
    PCA1 Avg stream size $\to$ Species richness & (+) & Post et al. (2000); Doi et al. (2009); McHugh et al. (2010) \\
    \hline
    PCA2 Avg temperature $\to$ Species richness & (+/-) & O’Gorman et al. (2017); Gauzens et al. (2020) \\
    \hline
    PCA1 Avg stream size $\to$ Connectance & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ Connectance & (+/-) & Petchey et al. (2010) \\
    \hline
    Species richness $\to$ Connectance & (-) & Winemiller (1989); Dunne (2006) \\
    \hline
    PCA1 Avg stream size $\to$ Avg trophic level & (+) & Post et al. (2000); Doi et al. (2009); McHugh et al. (2010) \\
    \hline
    PCA2 Avg temperature $\to$ Avg trophic level & (+/-) & O’Gorman et al. (2017)\\
    \hline
    Species richness $\to$ Avg trophic level & (+) & Schneider et al. (2016)\\
    \hline
    Species richness $\to$ Synchrony & (-) & Thébault and Loreau (2005); Loreau and de Mazancourt (2008); Thibaut and Connolly (2013) \\
    \hline
    Connectance $\to$ Synchrony & (+) & Teng and McCann (2004)\\
    \hline
    Avg trophic level $\to$ Synchrony & (+) & Raimondo et al. (2004)\\
    \hline
    PCA1 Avg stream size $\to$ Synchrony & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ Synchrony & (+) & Loreau and de Mazancourt (2008)\\
    \hline
    Species richness $\to$ CVsp & (+) & Tilman (1996) \\
    \hline
    Connectance $\to$ CVsp & (+) & Thébault and Loreau (2005)\\
    \hline
    Avg trophic level $\to$ CVsp & (-) & McCann and Rooney (2009); Shanafelt and Loreau (2018)\\
    \hline
    PCA1 Avg stream size $\to$ CVsp & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ CVsp & (+) & O’Gorman et al. (2017); Tabi et al. (2019)\\
    \hline
    CVsp $\to$ Biomass stability & (-) & Thibaut and Connolly (2013)\\
    \hline
    Synchrony $\to$ Biomass stability & (-) & Thibaut and Connolly (2013)\\
    \hline
    Species richness $\to$ Total biomass & (+) & Schneider et al. (2016)\\
    \hline
    Connectance $\to$ Total biomass & NA & Wang and Brose (2018)\\
    \hline
    Avg trophic level $\to$ Total biomass & (+) & Wang and Brose (2018)\\
    \hline
    PCA1 Avg stream size $\to$ Total biomass & (+) & Post et al. (2000); Allan and Ibanez Castillo (2009); Doi et al. (2009)\\
    \hline
    PCA2 Avg temperature $\to$ Total biomass & (+/-) & O’Gorman et al. (2017)\\
    \hline
    \end{tabular}
\end{center}
\end{table}

---
#### SEM hypothesis
#
#We linked environmental variables both to community structure and stability
#components. Rooted in assembly theory, environment filters species that can
#establish in a community and thus affect community structure. (+ add specific
#hypothesis about physical and climate). (1) We hypothesized that enrichment
#(characterized by DBO) and mean annual temperature will increase species turnover
#[@friedpetersen_drivers_2020]. (2) We hypothesized environmental forcing can also
#decrease biomass stability directly without going by community structure. First,
#instability of habitat and environmental factors may in turn drive biomass
#instability [@hansen_climate_2013]. Secondly, average environmental conditions
#such as enrichment can increase directly CV by increasing species biomass
#fluctuations [@tabi_warming_2019]. Increasing temperature was also hypothesized
#to decrease stability through starvation resulting from high metabolic activity
#[@tabi_warming_2019].
#
#---
## Species richness to community attributes 
#---
#
#To reveal the mechanisms behind the relationship between species richness and
#biomass stability, causal links were set from species richness to other
#community attributes. In particular, species richness decreased connectance
#following results from @thebault_stability_2010; @winemiller_must_1989. But
#theoretically, some ecological networks can show the inverse relationship
#[@winemiller_must_1989]. (6) We hypothesized also that species richness will
#modulate average weighted trophic level because of food web assembly processes.
#A community with more fish species can support higher trophic level. Finally, we
#set up causal relationships from community attributes to stability components. 
#
#---
##  Community attributes to stability
#---
#We hypothesized that (7) species richness decreased synchrony between species
#[@loreau_species_2008] and increased $\overline{CV_{sp}}$
#[@tilman_biodiversity_2006]. (8a) We expected that average trophic level had a
#positive or negative relationship with synchrony, bottom-up effects should
#result in positive whereas top-down in negative relationships. We expected a (8b)
#negative relationship between average trophic level and $\overline{CV_{sp}}$ if
#predators stabilised prey fluctuations or if the increasing part of
#biomass in the predators decreased $\overline{CV_{sp}}$ because of their higher
#body mass (if mean increase faster than variance). (9a) Connectance was expected
#to decrease or increase $\overline{CV_{sp}}$ and (9b) synchrony because of top
#down effects or bottom up effects.
---

```{r, cache = FALSE}
hist_var_caption <- paste0(
  "Distribution of the variables included Structural Equation Models The
  biomass stability, the CV of species (CVsp), PCA1, PCA2, species richness,
  synchrony and total biomass were logged."
)
```

```{r hist-sem-data, cache = FALSE, fig.cap = hist_var_caption}
sem_data %>%
  select(log_bm_std, log_rich_tot_std, log_sync, log_stab_std, log_cv_sp, log_RC1, log_RC2, t_lvl, ct) %>%
  gather(variable, value) %>%
  mutate(variable = str_replace_all(variable, get_sem_var_name_replacement())) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free")
```

```{r, cache = FALSE}
colin_sem_cap <- paste0(
  "Variance Inflation Factor for each individual linear model of the structural
  equation models."
) 
```

```{r s-colin-sem, fig.cap = ""}
mod_stab_list <-
  list(
    nlme::lme(log_rich_tot_std ~ log_RC1 + log_RC2,  random = ~ 1 | basin, data = sem_data),
    nlme::lme(ct ~ log_RC1 + log_RC2 + log_rich_tot_std, random = ~ 1 | basin, data = sem_data),
    nlme::lme(t_lvl ~ log_RC1 + log_RC2 + log_rich_tot_std, random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_sync ~ log_rich_tot_std + ct + t_lvl + log_RC1 + log_RC2,
      random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_cv_sp ~ log_rich_tot_std  + ct + t_lvl
      + log_RC1 + log_RC2, random = ~ 1 | basin, data = sem_data),
    lm(log_stab_std ~ log_cv_sp + log_sync, data = sem_data),
    nlme::lme(log_bm_std ~ log_rich_tot_std  + ct + t_lvl
      + log_RC1 + log_RC2, random = ~ 1 | basin, data = sem_data)
  )
names(mod_stab_list) <- c(
  "Species richness",
  "Connectance",
  "Avg trophic level",
  "Synchrony",
  "CVsp",
  "Biomass stability",
  "Total biomass"
)

colin_stab_sem <- map(mod_stab_list, car::vif)

tmp <- map(colin_stab_sem, enframe)

# colinearity table
colin_stab_sem <- tibble(vif = tmp, mod = names(mod_stab_list)) %>%
  unnest(vif) %>%
  rename(`Response variable` = mod) %>%
  mutate(
    value = round(value, 1),
    name = str_replace_all(name, get_sem_var_name_replacement())
    ) %>%
  spread(name, value)

```

```{r, cache = FALSE}
kable(colin_stab_sem,
  format = "latex",
  booktabs = T,
  caption = "Variance Inflation Factor for the linear models used in the structural
  equation models. NA values indicate that the term in column was not included
  in the modeling of the response variable in row."
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r, cache = FALSE}
caption_mod_res <- paste0(
  "Residuals versus fitted values of the linear models included in the structural
  equation models."
) 
```

```{r mod-res, fig.width = 7, fig.asp = 2, fig.cap = caption_mod_res, cache = FALSE}
library(nlme)
plot_residuals <- function (x, y = "Diagnostic") {
  plot.lme(x, main = y, xlab = "Fitted values", ylab = "Residuals")
}
p_resid_stab_sem <- map2(mod_stab_list, names(mod_stab_list) , plot_residuals)
plot_grid(plotlist = p_resid_stab_sem, ncol = 2)
```

```{r}
format_coef_table <- function (sem = NULL) {

  params <- sem$coefficients[, -length(colnames(sem$coefficients))]
  #colnames(params)[length(colnames(params))] <- "Stars"

  replace_wo_n <- str_replace_all(get_sem_var_name_replacement(), "\n", "")
  names(replace_wo_n) <- names(get_sem_var_name_replacement())

  params %<>%
    mutate_at(c("Response", "Predictor"),
      ~str_replace_all(., replace_wo_n))

  params
}
```

```{r}
est_stab_sem_caption <- paste0(
  "Estimate of the effects for the temporal stability of annual biomass of the structural equation
  model."

)

format_coef_table(stab_sem_rich) %>%
  mutate(Std.Estimate = cell_spec(Std.Estimate, bold = ifelse(P.Value <= 0.05, T, F), format = "latex", escape = F)) %>%
  kable(
    format = "latex",
    booktabs = T,
    caption = est_stab_sem_caption,
    escape = F 
    ) %>%
  collapse_rows(columns = 1, valign = "top") %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```

```{r}
est_bm_sem_caption <- paste0(
  "Estimate of the effects for the structural equation
  model of the median of annual total biomass."
)

format_coef_table(bm_sem_rich) %>%
  mutate(Std.Estimate = cell_spec(Std.Estimate, bold = ifelse(P.Value <= 0.05, T, F))) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = est_bm_sem_caption,
  escape = F 
  ) %>%
collapse_rows(columns = 1, valign = "top") %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r rsq-model}
rm_n <- str_remove_all(get_sem_var_name_replacement(), pattern = "\n")
names(rm_n) <- names(get_sem_var_name_replacement()) 
rsq_table <- map_dfr(mod_stab_list, piecewiseSEM::rsquared, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2)) %>%
  mutate(Response = str_replace_all(Response, rm_n)) %>%
  dplyr::select(Response, Marginal, Conditional, R.squared)
rsq_table_caption <- paste0(
  "R squared of the linear models used in the structural equation modeling."
)
```

```{r tab-coeff-bm, fig.cap = "Coefficient"}
# Replacement rules for term
rp_est <- c(
  "(Intercept)" =     "intercept",
  "sd\\_\\(intercept\\)\\.basin" = "sd intercept basin",
  "log_rich" =     "Median richness (log base 10)",
  "troph_group3" =     "high trophic group",
  "log_rich:troph_group3" =     "Median richness (log10) : high trophic group",
  "sig01.basin" =     "sd intercept basin",
  "sd_Observation.Residual" =     "residuals",
  "sigma" = "residuals"
)
# Replacement rules for group
rp_group <- c(
  "fixed" = "Fixed",
  "basin" = "Random"
)
```

```{r}
mask_drop_lm <- names(mod_stab_list) %in% "Biomass stability"
random_coeff_sem <- map(mod_stab_list[!mask_drop_lm], random.effects)
test <- getVarCov(mod_stab_list[!mask_drop_lm][[2]])

# get random effect 
get_intercept_variance <- function (mod) {
  vc <- VarCorr(mod)
  suppressWarnings(storage.mode(vc) <- "numeric")
  enframe(vc[1,])
}
random_coeff_sem <- map(mod_stab_list[!mask_drop_lm],
  get_intercept_variance) %>%
  enframe(name = "variable") %>%
  unnest(value) %>%
  pivot_wider(names_from = "name", values_from = "value")
#check
#all.equal(random_coeff_sem[[2]],(random_coeff_sem[[3]])^2)
#good

# get intercept 
get_intercept <- function (mod) {
  mod[["coefficients"]]$fixed["(Intercept)"]
}
intercept_sem <- map_dbl(
  mod_stab_list[!mask_drop_lm], get_intercept) %>%
  enframe(name = "variable", value = "Intercept")

formated_random_effects <- intercept_sem %>% 
  left_join(random_coeff_sem, by = "variable") %>%
  rename(`Response variable` = variable)
caption_random_effect <- paste0(
  "Estimated variance and standard deviation on the intercept due to the hydrographic basins."
)

kable(formated_random_effects,
  format = "latex",
  booktabs = T,
  caption = caption_random_effect
  ) %>%
kable_styling(latex_options = c("striped"))
```

```{r sensitivity, cache = TRUE, message = FALSE, fig.dim = c(10, 7), fig.cap = "Sensivity analysis with different dataset"}
library(furrr)

myload(op_analysis, op_analysis_wo_holes, op_analysis_ov, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)

myload(network_metrics, dir = mypath("data", "classes"))
sem_sensivity_data <- list(full = op_analysis, bbb = op_analysis_bbb,
  without_holes = op_analysis_wo_holes, bbb_without_holes = op_analysis_wo_holes_bbb)

sem_stab_l_data <- 
  furrr::future_map(sem_sensivity_data,
    ~build_dataset_get_sem_coefficient(
      .op = .x,
      sem_fun = compute_stab_sem_rich
    )
  )

nb_station <- map_int(sem_sensivity_data, ~length(unique(.x$station)))

sem_stab_results <- map(sem_stab_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_stab_results) <- str_to_lower(colnames(sem_stab_results))
mysave(sem_stab_results, dir = mypath("data"), overwrite = TRUE)
```

```{r sem-sens-bm, fig.cap = "Sensivity analysis for Biomass"}
sem_bm_l_data <- furrr::future_map(sem_sensivity_data,
  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_prod_sem_rich)
)
sem_bm_results <- map(sem_bm_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

# Reformat
colnames(sem_bm_results) <- str_to_lower(colnames(sem_bm_results))
mysave(sem_bm_results, dir = mypath("data"), overwrite = TRUE)
```

```{r}
sens_cap <- paste0(
  "Sensivity analysis of the Structural Equation models for total biomass
  and its temporal stability."," The Constant dataset is the one kepting only the
  sites that had not a significant temporal trend in biomass; ", "Constant/
    without holes dataset kept only the constant sites plus the ones that had not
  holes in their sampling period; ", "Complete is the dataset containing all the
  site that have been followed ten years or more, as the other dataset;", " Without holes dataset is
  the dataset containing only the sites that had no holes in the their sampling period."
)
```

```{r, fig.height = 8, fig.cap = sens_cap}
p_sens_bm <- plot_sensivity_analysis(
  .data = sem_bm_results,
  inverse_label_order = TRUE) +
  theme(legend.position = "none", plot.margin = unit(c(0, 0, 0, 25), units = "mm"))

p_sens_stab <- plot_sensivity_analysis(.data = sem_stab_results,
  inverse_label_order = TRUE) +
  guides(color = guide_legend(direction = "horizontal", nrow = 2))
legend_b <- get_legend(p_sens_stab)
p_sens_stab %<>% + theme(legend.position = "none", plot.margin = unit(c(0,0,0,27), units = "mm"))
p <- plot_grid(
  p_sens_bm, p_sens_stab, legend_b, nrow = 3,
  rel_heights = c(1, 1, .1),
  labels = c("Total biomass", "Biomass stability")
)
p
```

```{=latex}
\pagebreak
```

# References