### 2021-04-26 manuscript/bef_stability/table/hypothesis_sem.tex

\begin{tabular}{| m{7cm}| m{2cm}| m{4cm}|}
    \hline
    Link & hypothesis & reference\\
    \hline
    PCA1 Avg stream size $\to$ Species richness & (+) & \citet{post_ecosystem_2000,doi_resource_2009,mchugh_dual_2010} \\
    \hline
    PCA2 Avg temperature $\to$ Species richness & (+/-) & \citet{ogorman_unexpected_2017,gauzens_biodiversity_2020} \\
    \hline
    PCA1 Avg stream size $\to$ Connectance & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ Connectance & (+/-) & \citet{petchey_predicting_2010} \\
    \hline
    Species richness $\to$ Connectance & (-) & \citet{winemiller_must_1989, dunne_network_2006} \\
    \hline
    PCA1 Avg stream size $\to$ Avg trophic level & (+) &
    \citet{mchugh_dual_2010, doi_resource_2009, post_ecosystem_2000}\\
    \hline
    PCA2 Avg temperature $\to$ Avg trophic level & (+/-) & \citet{ogorman_unexpected_2017}\\
    \hline
    Species richness $\to$ Avg trophic level & (+) & \citet{schneider_animal_2016}\\
    \hline
    Species richness $\to$ Synchrony & (-) & \citet{thebault_trophic_2005, loreau_species_2008, thibaut_understanding_2013} \\
    \hline
    Connectance $\to$ Synchrony & (+) & \citet{teng_dynamics_2004}\\
    \hline
    Avg trophic level $\to$ Synchrony & (+) & \citet{raimondo_interspecific_2004}\\
    \hline
    PCA1 Avg stream size $\to$ Synchrony & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ Synchrony & (+) & \citet{loreau_species_2008}\\
    \hline
    Species richness $\to$ CVsp & (+) & \citep{tilman_biodiversity_1996} \\
    \hline
    Connectance $\to$ CVsp & (+) & \citep{thebault_trophic_2005}\\
    \hline
    Avg trophic level $\to$ CVsp & (-) & \citet{mccann_more_2009, shanafelt_stability_2018}\\
    \hline
    PCA1 Avg stream size $\to$ CVsp & (?) & \\
    \hline
    PCA2 Avg temperature $\to$ CVsp & (+) & \citep{ogorman_unexpected_2017, tabi_warming_2019}\\
    \hline
    CVsp $\to$ Biomass stability & (-) & \citet{thibaut_understanding_2013}\\
    \hline
    Synchrony $\to$ Biomass stability & (-) & \citet{thibaut_understanding_2013}\\
    \hline
    Species richness $\to$ Total biomass & (+) & \citet{schneider_animal_2016}\\
    \hline
    Connectance $\to$ Total biomass & NA & \citet{wang_biodiversity_2018}\\
    \hline
    Avg trophic level $\to$ Total biomass & (+) & \citet{wang_biodiversity_2018}\\
    \hline
    PCA1 Avg stream size $\to$ Total biomass & (+) & \citet{post_ecosystem_2000, doi_resource_2009, @allan_stream_2007}\\
    \hline
    PCA2 Avg temperature $\to$ Total biomass & (+/-) & \citep{ogorman_unexpected_2017}\\
    \hline
\end{tabular}