### 2021-07-06 manuscript/bef_stability/manuscript.Rmd
rewiew according to reviewer 1 and 2

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
  word_document:
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
nocite: |
    @declerck_species_2006
csl: ecology-letters.csl
---

```{r, echo = FALSE, results = FALSE, message = FALSE}
mypath <- rprojroot::find_package_root_file
data_common <- mypath("data")
dest_dir <- mypath("data", "species")

knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE 
)

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))

theme_set(theme_alain())
```

```{r, results = FALSE}
knitr::opts_chunk$set(
  results = TRUE 
)
```

- Author affiliation:
  - Alain Danet: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France
  - Maud Mouchet: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France
  - Willem Bonnaffé: Ecological and Evolutionary Dynamics Lab. Zoology Research
    and Administration Building, 11a Mansfield Rd, Oxford OX1 3SZ, United
    Kingdom.
  - Elisa Thébault: Sorbonne Université, CNRS, IRD, INRAE, Université Paris Est
    Créteil, Institute of Ecology and Environmental Sciences of Paris
    (iEES-Paris), Paris, France  
  - Colin Fontaine: Centre d’Ecologie et des Sciences de la Conservation, UMR 7204
    MNHN-CNRS-Sorbonne Université, Muséum national d’Histoire naturelle de
    Paris, 43 rue Buffon, 75005 Paris, France

- Author contributions: AD, CF, MM and ET designed the study. WB designed the
  network inference method and collected the food diet data. AD performed the
  data analysis and wrote the first draft of the manuscript. All authors
  contributed substantially to revisions.

- Data accessibility statement: Upon acceptance of the paper, we confirmed that
  the data supporting the results will be archived in an appropriate public
  repository and that the data DOI will be included at the
  end of the article. The code used for the analysis and to generate the
  manuscript is available on github: (alaindanet/fishcom.git). The raw fish
  database is public and available upon request to eddy.cosson@afbiodiversite.fr and
  thierry.point@afbiodiversite.fr. The raw data of water temperature, flow and
  Biological Oxygen Demand are available on http://www.naiades.eaufrance.fr/.

- Article type: Letters

- Running title: food-web structure, biomass and stability

- Number of words:
  - Abstract: 148
  - Main text: 4917

- Number of references: 102
- Number of figures: 3
- Number of tables: 1
- Number of text boxes: 0

- Coresponding author:
  - Alain Danet
  - UMR 7204 CESCO, CP 135; 43, rue Buffon, 75005 Paris, France
  - alain.danet@mnhn.fr 
  - +33140798134

# Abstract

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in
natural contexts. Here we estimated the links among environmental conditions,
richness, food-web structure, annual biomass and its temporal stability using a
standardised monitoring dataset of 99 stream fish communities spanning from 1995
to 2018. We first revealed that both richness and average trophic level are
positively related to annual biomass, with effects of similar strength. Second,
we found that community stability is fostered by mean trophic level, while
contrary to expectation, it is decreased by species richness. Finally, we found
that environmental conditions affect both biomass and its stability mainly via
effects on richness and network structure. Strikingly,
community stability was mediated by the variation of population stability rather
than synchrony, which contrasts with results from single-trophic communities. We
discuss the hypothesis that it could be a characteristic of multi-trophic
communities.

# Introduction

---
#### Biodiversity and ecosystem properties 
---

Current biodiversity losses and further global change have the potential to
affect ecosystems in complex ways [@loreau_biodiversity_2001;
@tilman_biodiversity_2014].  Biodiversity has indeed a major effect on ecosystem
functions and their stability in response to perturbations [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009]. Understanding the relationship between biodiversity and
ecosystem properties is thus crucial to better anticipate the impacts of global
change.

---
#### Species richness increases total biomass
---

Species richness has been found to increase primary productivity, community
biomass and its temporal stability in various ecosystems and taxa both
experimentally and in natural settings [@tilman_productivity_1996;
@cardinale_species_2002; @tilman_biodiversity_2006; @franssen_annual_2011;
@grace_integrative_2016; @pennekamp_biodiversity_2018; @houlahan_negative_2018;
@olivier_urbanization_2020]. However, these studies mainly focused on communities
composed of a single trophic level, generally primary producers, thereby
ignoring potential effects of food-web structure [but see
@scherber_bottom-up_2010].
Conversely, theoretical works suggest that both species richness and food-web
structure affect ecosystem properties such as productivity, total biomass and
its temporal stability. For example, the presence of species at higher trophic
levels may increase total primary production through top-down effects, which
result in the selection of primary producers that are larger and more
complementary in their use of resources [@wang_biodiversity_2018]. In communities
including producers and consumers, the food-web connectance is predicted to
modulate the diversity-productivity relationship [@thebault_food-web_2003;
@thebault_cascading_2007]. Similarly, theoretical studies generally show strong
links between food-web structure and stability [e.g.  @may_will_1972;
@dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011]. Overall, theoretical
models suggest that both species richness and food-web structure should affect
community biomass and its stability, but empirical evidence of such effects are
lacking.

While community stability is measured in many ways in theoretical studies, most
empirical studies measure community stability as the inverse of the coefficient
of variation of community productivity, biomass or abundance, also called
temporal stability [@kefi_advancing_2019]. This stability metric is also well
grounded in theory [e.g. @thebault_trophic_2005; @loreau_species_2008;
@thibaut_understanding_2013] and can be decomposed into two components:
population stability and synchrony [@thibaut_understanding_2013]. Studies
considering a single trophic level [@yachi_biodiversity_1999;
@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020] as
well as theoretical prediction for multitrophic communities
[@thebault_trophic_2005] tend to highlight that asynchrony in population
dynamics is key for the stabilising effects of species richness on community
biomass. Theory predicts that higher food-web connectance may increase
interaction strength and thereby increase population variability,
counterbalancing the positive effect of species richness on the temporal
stability of community biomass [@thebault_trophic_2005]. Predators can also have
cascading effects on both population variability and synchrony at lower trophic
levels [@teng_dynamics_2004; @shanafelt_stability_2018]. Overall, these
theoretical studies show that the effects of species richness and food-web
structure, such as connectance and average trophic level, are potentially
conflicting or synergistic in such a way that they need to be simultaneously
assessed.

The joint study of the effects of species richness and food-web structure on
community biomass and its stability is important in the context of environmental
changes. Recent studies comparing natural communities showed that environmental
characteristics determine in part community biomass [@grace_integrative_2016;
@woods_testing_2020] and its temporal stability [@franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2018;
@friedpetersen_drivers_2020]. The environment can affect ecosystem function or
stability directly. For instance, larger freshwater ecosystems are expected to
have more available resources and thus community biomass [@post_ecosystem_2000;
@doi_resource_2009] while agricultural intensification destabilises bird
communities by decreasing population stability [@olivier_urbanization_2020]. The
effects of environment can also be mediated by changes in species richness and
food-web structure. Larger ecosystems host food-webs with higher diversity and
trophic levels [e.g., @mchugh_dual_2010] and temperature is also known to affect
food-web diversity and structure  [e.g., @edeline_ecological_2013;
@ogorman_unexpected_2017; @gauzens_biodiversity_2020; @bonnaffe_trophic_2021].
Environmental characteristics have thus to be accounted for if we are to tease
apart the effects of species richness and food-web structure on community
biomass and its temporal stability.

Environment can affect functioning and stability directly or through effects on
resource availability, species richness or food-web
structure. Agricultural intensification and temperature changes can also...

---
#### Stream fish communities is a good model to study effects of horizontal & vertical diversity
---
Stream fish communities constitute a good model to study the effects of network
structure on community biomass and its temporal stability because their trophic
interactions are well-known. A substantial body of literature supports that
trophic interactions can be inferred through body size and diet information
[@jennings_weak_2001; @beckerman_foraging_2006; @gravel_inferring_2013;
@brose_predator_2019; @portalier_mechanics_2019]. Using long-term monitoring of stream fish
communities and inference of trophic interactions [@bonnaffe_trophic_2021], we
provide one of the first empirical tests of the expected relationship between
BEF and food-web complexity-stability accounting for environmental
characteristics.

# Methods

```{r, fig.show = "hide"}
myload(
  hab_analysis,
  temporal_station_desc,
  temporal_press_polluants,
  geo_station,
  dir = mypath("data")
)
habitat_press <- build_habitat_pressure_dataset(
  .habitat_analysis = hab_analysis,
  .tmp_press = temporal_press_polluants,
  .tmp_st_desc = temporal_station_desc,
  .geo_st = geo_station 
)

# Make PCA 
habitat_press %<>% na.omit()
mask <- names(habitat_press) %in% names(get_pca_var_name_replacement())

pca_rotated <- compute_rotated_pca(
  .data = habitat_press[, mask],
  naxis = 2
)
habitat_pressure <- 
  cbind(habitat_press[, c("station", "temperature_med", "alt")], pca_rotated$rotated$score) %>%
  as_tibble

mysave(habitat_press, habitat_pressure, dir = mypath("data"), overwrite = TRUE)
```

```{r}
prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))
tot_var_expl_pca <- sum(prop_var_expl_pca) 
```

```{r}
# Basin
st_basin <- get_basin_station(sf_obj = FALSE)

# Select fishing operation 
myload(op_analysis, op_analysis_wo_holes, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
range_surface <- map_dbl(c(min, max), ~.x(op_analysis_bbb$surface, na.rm =TRUE))
range_surface[1] <- ceiling(range_surface[1] / 10) * 10
range_surface[2] <- ceiling(range_surface[2] / 100) * 100
# quantile(op_analysis_bbb$surface, na.rm =TRUE)
```

## Data preparation and filtering

Fish communities were monitored across stream sections in metropolitan France
over the period 1995-2018 by the French Office of Water and Aquatic Ecosystems
(ONEMA) using electrofishing. Two different standardised protocols were used for
small and large streams (detailed in appendix B, part 1.1). In the case where a
stream was sampled alternatively with one of the two sampling protocols, we only
included data derived from the most frequently used sampling method. The
sampling season spanned from late spring to autumn. There was low heterogeneity
in the sampling month within stations, with the median sampling month being
mid-August and the median standard deviation of the sampling month was $0.7$ month,
i.e. three weeks.

We selected the stations that had been sampled ten years or more. As the concept
of temporal stability often referred to the variability around an equilibrium
point [@donohue_navigating_2016], we selected the stations that did not show temporal
trends in community biomass (detailed in appendix B, part 2). However, the
results were robust to the relaxation of this hypothesis (Fig. S7). These
selection steps resulted in 99 stations distributed over seven hydrographic
basins (Fig. \@ref(fig:matmet)A).

## Community biomass and its temporal stability

### Biomass estimation

The fishes collected by electrofishing were grouped in batches according to
their body size class and  species identity (Fig. S1, appendix B, part 1.1).
When individuals were too numerous in a given batch, the body size of a
subsample of the individuals was measured. The body size of the unmeasured
individuals was inferred under the assumption that the  body size of the
individuals in a batch follows a normal distribution (Fig. S1, appendix B,
part 1.2). The body mass of each individual was then estimated following the
results of two meta-analyses on fishes [@lleonart_removing_2000;
@froese_cube_2006], using an allometric law $B_i = 0.01 \times l_i^{3.03}$, with
$B$ the body mass in grams, $l$ the body size in millimetres and $i$ the
individual. Population biomass and community biomass at each sampling event were
then calculated by summing individual body masses. To account for the variation
in sampling effort among sites, population biomass and community biomass were
corrected by the surface sampled (range 
`r range_surface[1]` - `r range_surface[2]` square metres), and expressed in
grams per square metre ($gm^{-2}$). In the analysis, the community biomass of a
site was defined as the median of the community biomass across sampled years.

### Stability measures

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the community
biomass over time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$
being respectively the standard deviation and the temporal average of the annual
community biomass of the station $i$). Then, temporal stability ($S_i$) was
equal to $CV_i^{-1} = \mu_i \sigma_i^{-1}$. The variability of community
biomass ($CV_{com, i}$) can be decomposed into a synchrony ($\phi$) component
and a weighted average population variability ($\overline{CV_{sp, i}}$)
component according to @thibaut_understanding_2013:

\begin{equation} 
  CV_{com, i} = \overline{CV_{sp, i}} \times \sqrt{\phi_i}
  (\#eq:cvcom)
\end{equation}

To assess the mechanisms driving stability of community biomass, we thus in addition measured
the synchrony and the weighted average population variability $\overline{CV_{sp,
i}}$, the average CV at the population-level in the station $i$, weighted by the
relative biomass of the species populations, $B_{k,i}$.

Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
community biomass over time ($\sigma^2_{x_{T,i}}$) and the squared sum of the
standard deviation of population biomass over time ($\sigma_{x_{k,i}}$), $i$
being the station $i$ and $k$ the species $k$ [@loreau_species_2008]. Synchrony
varies from 0 to 1; 0 indicating complete asynchrony and 1 perfect synchrony.
Consequently, the lower is synchrony, the higher are the compensatory dynamics.

\begin{equation} 
\overline{CV_{sp, i}} = \sum_k{\frac{B_{k,i} {CV}_{k, i}}{\sum_k{B_{k,i}}}}
(\#eq:cvsp)
\end{equation}

\begin{equation} 
\phi_i = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sqrt{\sigma^2_{x_{k,i}}}})^2} = \frac{\sigma^2_{x_{T,i}}}{(\sum_k{\sigma_{x_{k,i}}})^2}
(\#eq:phi)
\end{equation}

## Species richness and food-web structure

### Species richness

Species richness in a given site was defined as the total number of species
present across all the sampling events. We controlled species richness for
sampling effort by dividing the total number of species by the total surface
sampled in the given station. Species richness was then expressed in number of
species per square metre sampled.

### Food-web inference

```{r}
myload(metaweb_analysis, dir= mypath("data"))
myload(trophic_class, dir= mypath("data", "classes"))
meta <- metaweb_analysis
#metaweb_analysis$resource
#meta$metaweb[meta$resource, meta$resource]
```

We determined the food-web structure in two steps [@bonnaffe_trophic_2021]: (1) we
built a metaweb describing the trophic interactions among all possible fish size
classes and species as well as with non-fish resource classes (Fig. S2), then
(2) we determined the food-web corresponding to each sampling event by
extracting from the metaweb the subset of fish species and size classes present
at a given site and sampling event (Fig. \@ref(fig:matmet)C and D).

The food-web was inferred from body size and ontogenic diet shift following
previous studies [@gravel_inferring_2013; @poisot_structure_2016;
@brose_predator_2019; @bonnaffe_trophic_2021]. The diet of stream fishes, as for
many organisms, shifts throughout an individual’s life, and thereby with body
size. For example, larvae and small juveniles feed on plankton while adults may
feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was divided into nine body size classes, each corresponding to a
trophic species (appendix B, Fig. S2). A trophic species constitutes a group of
individuals from the same species and size class, which are expected to share
the same sets of prey and predators. Additionally, we added seven resource nodes
present in the ontogenic food diet database: detritus, biofilm, phytoplankton,
zooplankton, macrophages, phytobenthos and zoobenthos (grey nodes in Fig.
\@ref(fig:matmet)C & D). The trophic species and the resources constituted the
412 nodes of the metaweb, i.e. 45 fish species divided into nine body size
classes plus seven resource nodes. The inference of trophic interactions was
shown to be more accurate when considering body size classes, i.e. trophic
species, than species [@jennings_weak_2001].

The feeding interactions of a trophic species were determined by both its
species identity and its body size. The ontogenic fish diet database was filled
according to fishbase [@froese_fishbase_2021] and literature (appendix A, Table
S2). Each fish species had two or three life stages which are delimited by its
body size range (appendix B, Fig. S2). The life stage of a fish trophic species
was determined by the centre of its body size range, hereafter midpoint
(appendix B, Fig. S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic species was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous trophic
species and every trophic species whose midpoint in body size range was included
in the predation window (appendix B, Fig. S2). The fish-resource trophic links
were set between the trophic species and the resources nodes according to the
food items of their size class. Lastly, the trophic links among resource nodes
were set according to the literature [@allan_stream_2007; @hart_handbook_2008].

### Food-web structure

For each food-web, corresponding to a particular site and year, we computed the
connectance and mean trophic level. Food-web connectance describes the level of
generalism in the network and is calculated as the actual number of interactions
divided by the number of interactions if all species interact ($C = L / N^2$, L
and N being resp. the number of links and of nodes). The trophic level of the
nodes was computed as one added to the average trophic level of its food items.
The average trophic level $\overline{T}$ was computed as the average trophic
level of the trophic species weighted by their biomass (excluding resource nodes
for which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i
\times (\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the
trophic level and the biomass of the node $i$. For each site, the median of
connectance and average trophic level over time was used in subsequent analysis
(Fig. \@ref(fig:matmet)C & D).

```{r, message = FALSE, results = FALSE}
# Com
com_data <- compute_community_temporal_analysis(.op = op_analysis_bbb,
 type_network_metrics = "classes"
)
# Compute sem dataset
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync_std"]],
  nmds = NULL,
  basin = st_basin
  )
mysave(sem_data, dir = mypath("data"), overwrite = TRUE)
```

## Environmental variables

We characterised the sites by their altitude, slope, distance from source,
stream width, depth and strahler order, i.e. its position in the stream network
hierarchy. We obtained data on Biological Oxygen Demand (BOD), which is often
related to enrichment and eutrophication in aquatic systems
[@mallin_factors_2006], water temperature, flow and from the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/). As data from naiades are not collected on the same
site as our fish sampling sites, we performed an interpolation of the water
temperature and flow based on the spatial stream network models
[@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure is
detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except for
water temperature for which records began in 2006. Then, we computed the average
and temporal CV values of water temperature, flow, BOD, river width and depth.
As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method on the
eigenvalues distribution (appendix B, Fig. S5). The two first principal
components explained `r tot_var_expl_pca`% of the variance (Fig.
\@ref(fig:matmet)B). A varimax axis rotation was performed to maximise the
representation of the environmental variables by the principal components
[@kaiser_varimax_1958; @revelle_psych_2019]. The first axis represented a
gradient from small to big streams, characterised by their width and depth, but
also a gradient of distance from source. We reversed the values
of the second axis, so that is represented a gradient of increasing average
water temperature, average BOD and decreasing altitude and slope. Further
description of the variables are provided in appendix B (part 1.4).

## Statistical analysis

We analysed the relationships among environment, species richness, food-web
structure, community biomass and its temporal stability with structural equation
models as they enable the assessment of potential direct and indirect
relationships among multiple variables [@grace_structural_2008;
@grace_causal_2014; @lefcheck_piecewisesem_2016; @grace_integrative_2016]. The
structural equation model explains the median annual community biomass as a
function of the environment, species richness and food-web structure
[@maureaud_biodiversityecosystem_2019; @woods_testing_2020]. We assumed that the
environment, species richness and food-web structure directly affect community
biomass (Fig. \@ref(fig:sem-bm)A), that species richness affects food-web
structure [@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010]
and that environment affects species richness and food-web structure.

We used the same model structure to assess the drivers of temporal stability of
community biomass (Fig. \@ref(fig:sem-stab)A). Instead of the annual community
biomass, we modelled the synchrony and population variability which in turn
fully determined the temporal stability of community biomass. We linearized the
mathematical relationship linking temporal stability of community biomass $S_i$
to $\overline{CV_{sp, i}}$ and $\phi$ by taking the log of temporal stability
such as

\begin{equation} 
    \log S_i = - \log \overline{CV_{sp, i}} - \frac{1}{2} \log \phi_i
  (\#eq:stab)
\end{equation}

The structural equation models were based on Gaussian linear mixed-effects
models. All variables except the connectance and average trophic level were
log-transformed to fulfil model requirements (see residual plots and data
transformation in Fig. S5 and S6, Table S5, appendix B). We set the seven
hydrographic basins as a random effect on the intercept to account for their
heterogeneity. We checked for the absence of multicollinearity with Variance
Inflation Factor (VIF, Table S6, appendix B). The slopes estimated by the
structural equation models and reported in the main text were standardised in
order to compare their magnitude ($r_\delta = \beta \frac{\sigma_x}{\sigma_y}$,
$\beta$ being the unstandardised slope, $\sigma_x$ and $\sigma_y$ being
respectively the standard deviation of the predictor variable $x$ and of the
response variable $y$). Therefore, the standardised coefficients expressed the
variation of $x$ and $y$ in standard deviation units. According to the equation
\@ref(eq:stab), the unstandardised coefficients of synchrony and population
variability determining
temporal stability of community biomass are respectively -0.5 and -1.0. When
standardised, the relative comparison of those coefficients comes down to a
comparison of the standard deviation of these two variables. The deviation from
unstandardised effect values (i.e. -0.5 and -1.0) indicates that one component of community stability
displays more variations than the other, and that the determinants of
the more variable component will be those affecting community
stability the most. The paths whose slope had a $P$ value inferior or equal to $0.05$
were reported in the main text, all the path values are presented in Table S7
and S8. Indirect effects on community biomass and its temporal stability were
computed by multiplying slopes along the path. Only significant direct slopes
were included in the calculation of indirect effects. Total effects were
computed as the sum of direct and indirect effects.

Environmental variables were interpolated with the `openStars` and `SSN` R
packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the nlme and piecewieseSEM R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and to generate the manuscript are accessible on github
(alaindanet/fishcom). The analyses were performed with R version 3.6.3
[@r_core_team_r_2020].

```{r get-data}
library(piecewiseSEM)
library(lme4)
library(nlme)

#ctrl <- lmeControl(opt='optim' relative)
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optimx'))
```

```{r}
# Compute SEMs
stab_sem_rich <- compute_stab_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))

bm_sem_rich <- compute_prod_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))

def_env_x_pos <- c(from = 0, to = 3)
asp <- 1.6

library(DiagrammeR)
meta_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability","figs", "meta_stab_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "stab_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)

meta_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "meta_bm_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  metamodel = TRUE,
  inverse_y_node_pose = TRUE
)
p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "figs", "bm_sem"),
  format = "pdf",
  width = 700*2,
  height = 700*2 * asp,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = def_env_x_pos,
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("log_RC1", "log_RC2"),
  inverse_y_node_pose = TRUE
)
```

```{r get-sem-coeff}
coef_stab_sem <- coef(stab_sem_rich) %>%
  select(- 9) %>%
  filter(P.Value <= 0.05)

coef_bm_sem <- coef(bm_sem_rich) %>%
  select(- 9) %>%
  filter(P.Value <= 0.05)
```

```{r get-bm-sem-single-effect}
bm_sem_coef <- filter(coef_bm_sem, P.Value <= 0.05)

rich_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_rich_tot_std")
rich_on_tlvl <- get_clean_sem_coef(sem = bm_sem_coef, resp = "t_lvl", pred = "log_rich_tot_std")
rich_on_ct <- get_clean_sem_coef(sem = bm_sem_coef, resp = "ct", pred = "log_rich_tot_std")
tlvl_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "t_lvl")
ct_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "ct")
RC1_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC1")
RC2_on_bm <- get_clean_sem_coef(sem = bm_sem_coef, resp = "log_bm_std", pred = "log_RC2")
```

```{r get-bm-tot-effect-env}
bm_all_params <- get_coefficient_piecewisesem(sem = bm_sem_rich, p_val_thl = NULL)

# effect of env on bm 
env_on_bm <- get_env_on_bm(fit = bm_all_params, p_val_thl = 0.05,
  type = "std")
# rich on bm 
rich_on_bm <- get_rich_on_bm(fit = bm_all_params,
  p_val_thl = 0.05, type = "std")
# direct
ct_on_bm_trick <- list(direct = c(ct = ct_on_bm), total = c(ct = ct_on_bm))
tlvl_on_bm_trick <- list(direct = c(t_lvl = tlvl_on_bm), total = c(t_lvl = tlvl_on_bm))

tmp_bm_indir_table <- map(list(env_on_bm, rich_on_bm, ct_on_bm_trick,
    tlvl_on_bm_trick),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```

```{r}
stab_all_params <- get_coefficient_piecewisesem(sem = stab_sem_rich, p_val_thl = NULL)
# effect of env on stab
env_on_stab <- get_env_on_stab(fit = stab_all_params, p_val_thl = 0.05,
  type = "std")
rich_on_stab <- get_rich_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
ct_on_stab <- get_ct_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
tlvl_on_stab <- get_tlvl_on_stab(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")

tmp_stab_indir_table <- map(list(env_on_stab, rich_on_stab, ct_on_stab, tlvl_on_stab),
  ~as_tibble(c(.x, list(variable = names(.x[[1]]))))
  ) %>%
bind_rows()
```
```{r get-sem-single-effect}
stab_sem_coef <- filter(coef_stab_sem, P.Value < 0.05)

sync_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_sync")
cv_sp_on_stab <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_stab_std", pred = "log_cv_sp")
RC1_on_sync <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_sync", pred = "log_RC1")
tlvl_on_cv_sp <- get_clean_sem_coef(sem = stab_sem_coef, resp = "log_cv_sp", pred = "t_lvl")
```

```{r get-sem-tot-effect-env}
# effect of env on sync and cv_sp
env_on_sync <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_sync", type = "std")
env_on_cv_sp <- get_env_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_cv_sp", type = "std")
# env on stab via network 
env_on_t_lvl <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "t_lvl", type = "std")
env_on_ct <- get_env_on_com(fit = stab_all_params,
  p_val_thl = 0.05,
  com = "ct", type = "std")
env_on_richness <- get_env_on_rich(fit = stab_all_params,
  p_val_thl = 0.05, type = "std")
```

# Results

The structural equation models explained a large part of the variation of
species richness, connectance, synchrony, $\overline{CV_{sp}}$ and community
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 = .11$, Fig. \@ref(fig:sem-bm)A and \@ref(fig:sem-stab)A).
We found a negative relationship between species richness and mean trophic level
($r_\delta = `r rich_on_tlvl`$, $r_\delta$ meaning standardised slope) and
connectance ($r_\delta = `r rich_on_ct`$, Fig. \@ref(fig:sem-bm)A and
\@ref(fig:sem-stab)A), meaning that a part of the effects of species richness on
community biomass and its temporal stability went through food-web structure.

## Species richness and average trophic level are positively correlated with community biomass

```{r get-sem-tot-effect-rich}
rich_on_sync <- get_rich_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_sync", type = "std")
rich_on_cv_sp <- get_rich_on_stab_comp(fit = stab_all_params,
  p_val_thl = 0.05,
  stab_comp = "log_cv_sp", type = "std")
```

Species richness and mean trophic level had a significant positive direct effect
on community biomass ($r_\delta = `r rich_on_bm$direct`$ and $r_\delta = 
`r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-bm)A, B, C, Table
\@ref(tab:indir-sem)),
while connectance had no direct effect.  The two PCA axes synthesising
environmental variables had no significant direct effects on community biomass,
but indirect ones (Table \@ref(tab:indir-sem)).  The first axis, related to the
stream size, was positively linked to community biomass ($r_\delta = 
`r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)), through both
species
richness and mean trophic level 
($r_\delta = `r env_on_bm$via_richness["log_RC1"]`$ and 
$`r env_on_bm$via_tlvl["log_RC1"]`$
resp., Fig. \@ref(fig:sem-bm) F, G, Table \@ref(tab:indir-sem)). The second
axis, related to the average annual temperature and BOD was positively related
to community biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)) mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Fig. \@ref(fig:sem-bm) G, Table
\@ref(tab:indir-sem)). Warmer and enriched, i.e.  with higher BOD, streams
contained more species, and thus indirectly increased the biomass of the
communities. The total effect of PCA1, PCA2 and species richness on community
biomass were of the same magnitude.

## Temporal stability of annual community biomass

By its effects on synchrony and population variability ($\overline{CV_{sp}}$),
species richness had a negative total effect on biomass stability ($r_\delta =
`r rich_on_stab$total`$, Table \@ref(tab:indir-sem)). This effect resulted from
two opposite effects. On the one hand, species richness had a direct negative
effect on synchrony ($r_\delta = `r rich_on_sync$direct`$, Fig.
\@ref(fig:sem-stab)A, B), thereby increasing stability ($r_\delta = 
`r rich_on_stab$via_sync`$).  On the other hand, species richness had a positive
direct effect on $\overline{CV_{sp}}$ ($r_\delta = `r rich_on_cv_sp$direct`$,
Fig. \@ref(fig:sem-stab)A, C), thereby decreasing biomass stability ($r_\delta =
`r rich_on_stab$via_cv_sp`$, Table \@ref(tab:indir-sem)), which outweighed the
stabilising effect of species richness via decreased synchrony. The negative
effect of $\overline{CV_{sp}}$ on temporal stability was twice higher than the negative effect of
synchrony (resp.  $r_\delta = `r cv_sp_on_stab`$ and $r_\delta = 
`r sync_on_stab`$, Fig.  \@ref(fig:sem-stab)A), which is as expected from the mathematical
decomposition of the stability of community biomass. Population variability and
synchrony indeed contribute respectively to -1 and -0.5 when the coefficients
are not standardised (Eq. 4). Our result thus also means that the standard
deviation of $\overline{CV_{sp}}$ is of roughly the same magnitude as the one of
synchrony in this case.

The total effect of mean trophic level and species richness on temporal
stability of community biomass were of the same magnitude (resp. $`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)). The mean trophic level had only a direct negative effect
on $\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$, Fig.  \@ref(fig:sem-stab)A, F) and thereby
had a total positive effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$, Table \@ref(tab:indir-sem)). Food-webs in
which community biomass is located on average at higher trophic levels thus tend
to have higher stability of population and community biomass. Connectance had
no significant direct effect on either $\overline{CV_{sp}}$ nor synchrony and
thus no effect on temporal stability of community biomass (Fig.
\@ref(fig:sem-stab)A, Table \@ref(tab:indir-sem)).

The total effect of PCA2, related to average annual temperature and BOD, on
temporal stability was of higher magnitude than the effects of species richness
and mean trophic level (Table \@ref(tab:indir-sem)). The PCA2 had a total
negative effect on the temporal stability of communities, meaning that higher
temperature and BOD results in lower stability ($r_\delta = 
`r env_on_stab$total["log_RC2"]`$, Table \@ref(tab:indir-sem)). PCA2 had an
indirect positive effect on temporal stability through a direct positive effect
on mean trophic level ($r_\delta = `r env_on_stab$via_tlvl["log_RC2"]`$, Fig.
\@ref(fig:sem-stab) I, Table \@ref(tab:indir-sem)). However, PCA2 had a negative
effect on stability of community biomass through a direct positive effect on
$\overline{CV_{sp}}$ and species richness (resp.  $r_\delta = 
`r env_on_stab$via_cv_sp["log_RC2"]`$ and $r_\delta = 
`r env_on_stab$via_richness["log_RC2"]`$, Fig. \@ref(fig:sem-stab)G, Fig.
\@ref(fig:sem-bm)G, Table \@ref(tab:indir-sem)). PCA1, related to stream size,
had in contrast a very slight positive effect on stability of community biomass
($r_\delta = `r env_on_stab$total["log_RC1"]`$, Table \@ref(tab:indir-sem)).
Warmer and larger streams tend to have more species, but the variability of
their population biomass is respectively higher and lower, resulting in
respectively lower and higher temporal stability of community biomass.

```{r, cache = FALSE}
cap_sem <- paste0(
  "Direct, indirect and total effects of environment, species richness,
  connectance and mean trophic level on community biomass
  and its temporal stability. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable.
  Indirect effects are considered as one step from the target variable. For
  exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.62$. Only the
  significant direct effects were retained for the computation of indirect
  effects. ", "As an example, the indirect effect via species richness takes
  into account the direct effect of species richness as well as the effect of
  species richness through connectance and average trophic level. ",
 "NA values indicate that the corresponding paths do not exist in the
 specified Structural Equation Model (SEM). ", "NS: statistically non-significant."
)
```

```{r indir-stab, cache = FALSE}

full_table <- bind_rows(
  tmp_bm_indir_table,
  tmp_stab_indir_table
) %>%
mutate(`SEM type` = c(rep("Total annual biomass", 5), rep("Temporal Stability of annual biomass", 5))) %>%
select(`SEM type`, variable, direct, via_richness, via_ct, via_tlvl, via_cv_sp, via_sync, total) %>%
format_table() %>%
select(`SEM type`, everything())

full_table %<>%
  mutate(Direct = ifelse(Direct == 0, "NS", Direct)) %>%
  mutate_if(is.numeric, format, digits = 2, nsmall = 2)

library(kableExtra)
mytable <- kable(full_table,
  format = "latex",
  booktabs = T,
  label = paste0("indir-sem"),
  caption = cap_sem
  ) %>%
kable_styling(latex_options =c("scale_down")) %>%
add_header_above(c(" " = 3, "Indirect effects via" = 5)) %>%
column_spec(1:2, width = "3cm") %>%
collapse_rows(columns = 1, valign = "top")
#write(mytable, file = mypath("manuscript", "bef_stability", "table.tex"))
```

# Discussion

---
#This study is unique by its level of integration, from environmental
#effects to total biomass and its stability, going through species richness and
#food-web structure. It is the first direct assessment of a possible joint effect
#of species richness and network structure on both total biomass and its temporal
#stability. We found that species richness and mean trophic level had a positive
#effect on total biomass. In contrast, species richness had a negative effect on
#temporal stability whereas mean trophic level covaried positively. We also found
#that the effect of species richness and mean trophic level on total biomass and
#its temporal stability were not independent as species richness had a negative
#effect on mean trophic level. Furthermore, we found that connectance had no
#significant effect on both total biomass and its temporal stability. The results
#shows that temporal stability was mediated by changes in species variability 
#rather than synchrony. Finally, we found that the magnitude of the effects of
#average temperature and altitude was equal or superior to the effects of species
#richness and mean trophic level. 
---

## Species richness increased community biomass but decreased its temporal stability

We found that species richness increased community biomass in line with the
results in the plants and tree communities [@hector_plant_1999;
@tilman_biodiversity_2014; @wu_relationship_2015; @li_relationship_2018] as well
as those found for fish communities [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020] and for theoretical food-web models
[@schneider_animal_2016]. Theory suggests that the positive effect of species
richness on community biomass might be induced by resource use complementarity
[@carey_determining_2011], which is driven by niche differentiation
[@loreau_biodiversity_1998; @loreau_partitioning_2001] or driven by the
dominance of a highly productive species [i.e. positive selection effect,
@loreau_biodiversity_2001]. Recent findings suggest that such positive selection
effect [@loreau_partitioning_2001] is the main driver of the observed positive
relationship between biomass and species richness in fish communities
[@maureaud_biodiversityecosystem_2019; @woods_testing_2020].

Contrary to most experimental and empirical results [@tilman_biodiversity_2006;
@franssen_annual_2011; @pennekamp_biodiversity_2018; @olivier_urbanization_2020;
@valencia_synchrony_2020, but see, Declerck *et al.* 2006], we found a negative
relationship between species richness and stability of community biomass. This
effect resulted from a combination of two effects, one on population synchrony
and one on population variability, both of which contribute to the stability of
community biomass. The negative effect of species richness on population
synchrony we found, thereby increasing stability of community biomass, is in
accordance with most studies on experimental grasslands
[@isbell_biodiversity_2009; @roscher_identifying_2011] as well as natural
communities such as birds, grasses, bats or butterflies [@bluthgen_land_2016;
@olivier_urbanization_2020, but see, Declerck *et al.* 2006;
@valencia_synchrony_2020]. Our result thus extend this relationship to
multi-trophic communities.

The positive effect of species richness on population variability that we found,
thereby destabilising communities, is more intriguing and does not fit the
findings of a meta-analysis reporting more frequent negative effects of richness
on population variability in multi-trophic communities [@jiang_different_2009].
Food-web theory predicts that the relationship between the temporal stability of
populations and species richness depends on the strength of interactions and its
distribution [@mccann_reevaluating_1997; @thebault_trophic_2005;
@downing_temporal_2020], with stronger interactions promoting a negative
relationship between species richness and stability. This might be the case in
our study system but this hypothesis remains to be directly tested.
Alternatively, a positive richness-population variability relationship can come
from an increased demographic stochasticity with increasing species richness
[@loreau_species_2008].

In our study system, the destabilising effect of species richness via increased
population variability was stronger than the stabilising effect via decreased
synchrony. The balance between both effects is governed by the balance in the
variation in population variability and synchrony among communities, this
because we used standardised coefficients (see methods). Such an effect of
population variability on community stability twice as strong as the one of
synchrony is expected from the mathematical relation between these three
variables. This indicates that the variability of population variability and
synchrony among the studied communities are equivalent (Eq. \@ref(eq:stab)).
This result contrasts with previous findings on terrestrial animal communities
showing a variability in population synchrony twice as strong as the one of
population variability [@olivier_urbanization_2020]. Such difference in the
relative variability of population stability and synchrony between studies might
be related to the multitrophic structure of the dataset as theory predicts
potential synchronising effect of generalist predators, such as piscivorous
fishes, on prey dynamics [@raimondo_interspecific_2004; @huber_role_2006].
Trophic interactions might also raise synchrony between predators and preys and
thereby increase the overall synchrony of trophic communities [Fig. S9,
@bulmer_phase_1975]. However, generalist predators can also generate
asynchronous dynamics among preys depending on the distribution of interaction
strengths [@huber_role_2006; @mccann_more_2009]. The difference between our
results and those of @olivier_urbanization_2020 might also relate to the type of
ecosystems considered. Aquatic ecosystems have been suggested to show greater
population variability than terrestrial ecosystems [@rip_cross-ecosystem_2011],
which could also translate in greater variation range of population variability
among aquatic communities than among terrestrial ones. Another hypothesis
relates to the evenness of the biomass distribution among species, with low
evenness decreasing the range of variation in synchrony among communities and thereby
the strength of the portfolio effect [@doak_statistical_1998;
@thibaut_understanding_2013]. Whether the biomass distribution in stream fish
assemblages is particularly uneven when compared to the terrestrial communities
studied by @olivier_urbanization_2020 remains however to be tested. In any case,
more studies are needed to understand how the relative importance of synchrony
and population variability for community stability varies among ecosystems.

---
#@thibaut_understanding_2013 proposed that a
#strongly uneven biomass distribution across populations and that the biomass
#variance of population increases much faster than its mean can generate a
#positive effect of species richness on population ($\overline{CV_{sp}}$)
#stability.
---

## Average trophic level increased community biomass and its temporal stability while connectance had no effect

Our analysis further indicates that network structure, and in particular average
trophic level was strongly related to community biomass, as higher community
biomass was found in communities with higher average trophic level. These
results are in accordance with the theoretical predictions [@schneider_animal_2016;
@wang_biodiversity_2018], proposing that predation selects for primary producers with
higher resource use efficiency and enhance complementarity between primary
producers [@wang_biodiversity_2018].

Further, our results show that communities with higher average trophic level had
a lower population variability, and were thereby more stable. This result is
also in accordance with theoretical models predicting a higher stability of
populations of higher trophic levels [@shanafelt_stability_2018;
@barbier_pyramids_2019] or a damping effect of higher trophic levels on the
variability of lower trophic level populations
[@mccann_diversitystability_2000].

In contrast, we found that food-web connectance was not related to community
biomass, which contradicts theoretical models suggesting that higher connectance
is expected to lower community biomass by decreasing complementarity in resource
use among species [@thebault_food-web_2003; @poisot_trophic_2013]. However, the
effect of connectance on productivity remains poorly investigated.

Unexpectedly, we found no relationship between food-web connectance and
stability of community biomass which contrasts with many theoretical results
highlighting a food-web complexity-stability relationships [e.g.,
@angelis_stability_1975; @may_will_1972; @thebault_stability_2010; but see
@neutel_reconciling_2007], although they generally consider different stability
concepts and metrics [@arnoldi_resilience_2016]. The lack of connectance effect
on the temporal stability of community biomass might be related to the
connectance metric we used, in particular the fact that it did not account for
interaction strengths. Indeed, connectance stability relationship is expected to
vary depending on the strength of trophic interactions [@thebault_trophic_2005].
Further, in our dataset, resource nodes, i.e. primary producers, decomposers and
the benthic species, were highly aggregated which might blur potential
connectance differences among communities.

## Environment had an strong effect on community biomass and its temporal stability

We found that the larger as well as warmer and more enriched streams had higher
community biomass, and that this was mediated by higher species richness and
average trophic level. Regarding the effect of ecosystem size, our results thus
support a classical theory of food-web ecology [@post_ecosystem_2000;
@doi_resource_2009]. Larger freshwater ecosystems provide higher diversity of
resources through habitat heterogeneity and then result in higher species
richness and mean trophic level per surface unit, through complementarity in
resource use [@post_ecosystem_2000; @doi_resource_2009; @mchugh_dual_2010].
Regarding temperature and enrichment effect, our study joins the mounting body
of works reporting effect on the structure and biomass of aquatic systems
[@allan_stream_2007; @edeline_ecological_2013; @emmrich_geographical_2014;
@hattab_forecasting_2016; @woods_testing_2020; @bonnaffe_trophic_2021]. However,
current results do not show consensus [@yvondurocher_warming_2011;
@dossena_warming_2012; @maureaud_biodiversityecosystem_2019;
@tabi_warming_2019], with potential antagonistic effects of temperature and
enrichment as well as non-monotonic effects [@oconnor_warming_2009;
@uszko_effects_2017; @tabi_warming_2019]. Although we could not tease apart the
respective effects of temperature and enrichment, our results indicates that in
the range of observed covariation in temperature and enrichment, their joint
effects result in a positive relationship with community biomass.

Our analysis further revealed that stream size had only a weak effect on the
temporal stability of community biomass, while warmer and enriched streams
exhibit less stable community biomass mainly via an increase in population
variability. These results are in line with previous experiments showing that
warming or enrichment decrease the temporal stability of communities
[@kratina_warming_2012; @tabi_warming_2019]. However, several studies have
highlighted that the effects of temperature on stability can be complex, and
interact with those of enrichment [@binzer_dynamics_2012; @uszko_effects_2017].

## Network reconstruction: limits and perspectives

The lack of experimental or empirical studies considering the effects of both
species richness and food-web structure on ecosystem functions can be explained
by the challenges associated with manipulating both species richness and
food-web structure experimentally [e.g. @sander_ecological_2017] and by the fact
that describing trophic interactions between species is highly labour-intensive.
These challenges are even more acute when studying the temporal stability of
ecosystem properties, which requires time series. In this context, recent
developments to predict trophic interactions among species
[@beckerman_foraging_2006; @gravel_inferring_2013; @portalier_mechanics_2019],
combined with observational long-term monitoring of ecological communities,
offer a promising opportunity.  They offer a way to maximise the amount of
information that can be extracted from already collected empirical data, but at
the cost of their precision. Here, we built binary interaction networks based on
the ratio of body-size between predator and prey ignoring modulation of
interactions strength depending on preferences or environmental factors.
Refining the construction of interaction networks by, for example, integrating
predator behaviour [@portalier_mechanics_2019] or developing more statistically
flexible framework such as machine learning [@pichler_machine_2020] will
certainly increase the accuracy of species interaction predictions. Although
challenging, methods of food-web reconstruction have a great potential to
increase our understanding of network structure effects on ecosystem function
and dynamic in natural settings using available long-term empirical data.

# Conclusion

Our work contributes to bridging the gap between theoretical, experimental and
empirical findings on the functioning and stability of communities. Our study
extends the assessment of the effect of species richness on community biomass
and its temporal stability to the case of freshwater food-webs in natural
settings. While our results are in accordance with numbers of studies showing a
positive effect of richness on community biomass, they provide a contrasting
example of negative relationship between richness and stability of community
biomass. Our results highlight population variability as key in transmitting
effects of species richness, network structure and environmental gradients on
community stability, opening stimulating questions about how the food-web
structure and species richness drive population synchrony and variability.

```{r}
save.image(file = mypath("manuscript", "bef_stability", "result", "workspace.rda"))
```

# References {-}

<div id="refs"></div>

\clearpage

# Figures and Table {-}

```{=latex}
\input{table.tex}
```

\clearpage

```{r, cache = FALSE}
pca_caption <- paste0(
  "(B) Rotated axis of PCA on environmental variables. The five components
  explained, ", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  " Axis 1 was positively related to the stream size, axis 2 was related to
  the stream slope and altitude and negatively related to average water temperature and
  BOD. For simplicity, we reversed the second axis of the PCA in the structural equation
  models, so that the second axis covaries with average water temperature and
  BOD."
  )

matmet_cap <- paste0( 
  "(A) Observation sites (black dots) were located across France and monitored
  on the period 1995-2018. The red dots show the location of (C) and (D) sites.
  Inner borders draw the limits of the hydrographic basins. (B) Principal
  Component Analysis performed on environmental variables.  The first axis was
  positively related to the stream size, the second axis was related to the
  stream slope and altitude and negatively related to average water temperature
  and Biological Oxygen Demand (BOD). (C) and (D) Temporal dynamic of
  communities with examples of networks corresponding to the years 1997, 1999
  and 2002. Each trophic species is represented by a node and each color
  represents a species (see method). Node size is proportional to the biomass of
  the trophic species. The seven resource nodes are coloured in grey. Species
  names corresponding to the three digit codes are found in Table S1.")

```

```{r matmet, cache = FALSE, fig.cap = matmet_cap, results = "fig.show", out.height="80%"}
library(knitr)
include_graphics(path = mypath("manuscript", "bef_stability", "figs", "fig1_modif.pdf"))
```

\clearpage

```{r}

sem_bm_caption <- paste0( 
  "Relationships among the predictors of community biomass. (A) Structural
  Equation Model for the median of annual community biomass (in grams by square
    metre). Dotted and solid arrows represent, respectively, negative and
  positive path coefficients. Non significant paths are coloured in grey. The
  path coefficients are standardised. $R^2$: adjusted $R^2$; Avg: Average, CVsp: weighted
  temporal coefficient of variation of the population biomass; PCA1 and PCA2:
  respectively the first and second axis of the principal component analysis
  performed on environmental variables (see methods). (B) to (G): significant
  bivariate relationships between variables of the structural equation model.
  Solid lines depict the predictions of the statistical model. A sensibility
  analysis is displayed in supplementary information (Fig.
    S8).")

```

```{r sem-bm, out.width="\\maxwidth", fig.cap = sem_bm_caption, results = "fig.show", cache = FALSE}
#out.height="120%",
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple

#plot_grid(
  #meta_bm_sem_rich, p_bm_sem_rich,
  #meta_stab_sem_rich, p_stab_sem_rich,
  #labels = LETTERS[1:4], nrow = 2,
  #rel_widths = c(1, 1)
#)
# Put the panel of Colin
arrow_dir <- mypath("manuscript", "bef_stability","fleches", "figure_sem2")
filename <- list.files(arrow_dir, full.names = TRUE)
mask <- str_match(filename, "\\.svg")[, 1] %>% is.na %>% !.
#filename[mask]
type <- c("bm", "meta_bm", "meta_stab", "stab")
sem_list <- map(filename[mask], function (x) {
  w <- 2400
  img <- magick::image_read_svg(
    path = x, width = w, height=w*1.618)#, density = 600
  p_sem <- cowplot::ggdraw() +
    cowplot::draw_image(img)#, scale = 1.0
  return(p_sem)
})
sem_list %<>% map(., ~.x + theme(plot.margin = unit(x = rep(0, 4), units = "pt")))
names(sem_list) <- type
sem_plot <- plot_grid(
  plotlist = sem_list[c("meta_bm", "bm", "meta_stab", "stab")],
  labels = "AUTO" 
  )
save_plot(
  filename = mypath("manuscript", "bef_stability", "figs",
    "sem_colin.pdf"),
  plot = sem_plot, ncol = 2, nrow = 2,
  base_width = 2.71, base_asp = 1.2 
)
#sem_plot
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "fig2.pdf"))
```

\clearpage

```{r}

sem_stab_caption <- paste0(
  "Relationships among the predictors of the temporal stability of community
  biomass. (A) Structural Equation Model for the temporal stability of community
  biomass. Dotted and solid arrows represent, respectively, negative and positive
  path coefficients. Non significant paths are coloured in grey. The path
  coefficients are standardised. $R^2$: adjusted $R^2$; Avg: Average, CVsp: weighted
  temporal coefficient of variation of the population biomass; PCA1 and PCA2:
  respectively the first and second axis of the principal component analysis
  performed on environmental variables (see methods). The $R^2$ for the temporal
  stability of community biomass is exactly one because the relation is
  determistic (see Eq. 4). (B) to (I): significant
  bivariate relationships between variables of the structural equation model.
  Solid lines depict the predictions of the statistical model. A sensibility
  analysis is displayed in supplementary information (Fig.
    S8)."
  )
```

```{r sem-stab, out.width="\\maxwidth", fig.cap = sem_stab_caption, results = "fig.show", cache = FALSE}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "fig3.pdf"))
```